---
title: "Tidy Template"
author: "NearAndDistant"
date: "29/09/2021"
output: html_document
---

##  The Regression of Poverty

all data from, here: https://data.oecd.org/society.htm#profile-Inequality

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Setup & Import
```{r}
library(tidyverse)

# create driectory for new project
dir.create(here::here("the_regression_of_poverty"))
dir.create(here::here("the_regression_of_poverty/oecd_data"))

# import data for project
```


```{r}

# poverty rate, 18-65 year-olds, Ratio, 2011 – 2019, 
# source: https://data.oecd.org/inequality/poverty-rate.htm
pov_raw        <- read_csv("the_regression_of_poverty/oecd_data/DP_LIVE_10052022230151283.csv") %>% janitor::clean_names() %>% select(-measure, -frequency, -flag_codes, -subject)

# income inequality, gini coefficient, 0 = complete equality; 1 = complete inequality, 2011 – 2019, 
# source: https://data.oecd.org/inequality/income-inequality.htm#indicator-chart
gini_raw       <- read_csv("the_regression_of_poverty/oecd_data/DP_LIVE_10052022231527801.csv") %>% janitor::clean_names() %>% select(-measure, -frequency, -flag_codes, -subject)

# gross national income, total, US dollars/capita, 2011 – 2019
# source: https://data.oecd.org/natincome/gross-national-income.htm
gni_raw        <- read_csv("the_regression_of_poverty/oecd_data/DP_LIVE_10052022231846301.csv") %>% janitor::clean_names() %>% select(-measure, -frequency, -flag_codes, -subject)

# social spending, public, % of GDP, 2011 – 2019
# source: https://data.oecd.org/socialexp/social-spending.htm
pub_spend_raw  <- read_csv("the_regression_of_poverty/oecd_data/DP_LIVE_10052022232254535.csv") %>% janitor::clean_names() %>% select(-measure, -frequency, -flag_codes, -subject)

# Household disposable income, gross, per capita, percentage change, previous period, 2011 – 2019
# source: https://data.oecd.org/hha/household-disposable-income.htm
#dis_inc_raw    <- read_csv("the_regression_of_poverty/oecd_data/DP_LIVE_10052022232127910.csv") %>% janitor::clean_names() %>% select(-measure, -frequency, -flag_codes, -subject)

# working age population, total, % of population, 2011 – 2019
# source: https://data.oecd.org/pop/working-age-population.htm#indicator-chart
wrk_age_pc_raw <- read_csv("the_regression_of_poverty/oecd_data/DP_LIVE_10052022232724973.csv") %>% janitor::clean_names() %>% select(-measure, -frequency, -flag_codes, -subject)

# adult education level, tertiary, % of 25-64 year-olds, 2011 – 2019
# source: https://data.oecd.org/eduatt/adult-education-level.htm
adl_pc_edu_raw <- read_csv("the_regression_of_poverty/oecd_data/DP_LIVE_10052022232527613.csv") %>% janitor::clean_names() %>% select(-measure, -frequency, -flag_codes, -subject)

```

#### Clean & Wrangle
```{r}
library(countrycode)

# not in
`%!in%` <- Negate(`%in%`)

pov_indicators <-
  pov_raw %>%
  bind_rows(gini_raw, gni_raw, pub_spend_raw, wrk_age_pc_raw, adl_pc_edu_raw) %>%
  filter(location %!in% c("OECD", "OAVG")) %>%
  mutate(location  = countrycode(location, "iso3c", "country.name")) %>%
  mutate(indicator = case_when(indicator == "POVERTY"    ~ "pc_poverty",
                               indicator == "EDUADULT"   ~ "pc_higher_edu",
                               indicator == "GNI"        ~ "gni_per_capita",
                               indicator == "INCOMEINEQ" ~ "gini_inequality",
                               indicator == "SOCEXP"     ~ "pc_public_exp_of_gdp",
                               indicator == "WKGPOP"     ~ "pc_working_age_pop"))

# year filters
pov_indicators_2018 <-
  pov_indicators %>%
  filter(time == 2018) %>%
  select(-time) %>%
  pivot_wider(id_cols = location, names_from = "indicator", values_from = "value")  %>%
  mutate(across(.cols = c(pc_public_exp_of_gdp:pc_higher_edu), ~ .x / 100))

```

#### Analysis

```{r}
library(gridExtra)

dist_pov <-
pov_indicators_2018 %>%
  ggplot(aes(pc_poverty)) +
  geom_histogram(fill = "#1F456E" , color = "white") +
  labs(
    x = "Poverty Rate (%)",
    y = "Number of Counties"
  )

dist_edu <-
pov_indicators_2018 %>%
  ggplot(aes(pc_higher_edu)) +
  geom_histogram(fill = "#1F456E" , color = "white", na.rm = TRUE) +
  labs(
    x = "Higher Education Rate (%)",
    y = "Number of Counties"
  )

dist_gni <-
pov_indicators_2018 %>%
  ggplot(aes(gni_per_capita)) +
  geom_histogram(fill = "#1F456E" , color = "white", na.rm = TRUE) +
  labs(
    x = "GNI / Capita (USD)",
    y = "Number of Counties"
  )

#dist_hhdi <-
#pov_indicators_2018 %>%
#  ggplot(aes(household_disp_inc)) +
#  geom_histogram(fill = "#1F456E" , color = "white", na.rm = TRUE) +
#  labs(
#    x = "Household Disposable Income / Capita (Annual Chnage %)",
#    y = "Number of Counties"
#  )

# removed as too many missing datapoints

dist_gini <-
pov_indicators_2018 %>%
  ggplot(aes(gini_inequality)) +
  geom_histogram(fill = "#1F456E" , color = "white", na.rm = TRUE) +
  labs(
    x = "Gini Coefficient (0 = complete equality; 1 = complete inequality)",
    y = "Number of Counties"
  )

dist_pub <-
pov_indicators_2018 %>%
  ggplot(aes(pc_public_exp_of_gdp)) +
  geom_histogram(fill = "#1F456E" , color = "white", na.rm = TRUE) +
  labs(
    x = "Public Expenditure / GDP (%)",
    y = "Number of Counties"
  )

dist_wrk_age <-
pov_indicators_2018 %>%
  ggplot(aes(pc_working_age_pop)) +
  geom_histogram(fill = "#1F456E" , color = "white", na.rm = TRUE) +
  labs(
    x = "Working Age Population (% total pop)",
    y = "Number of Counties"
  )


# arrange
gridExtra::grid.arrange(dist_pov , dist_edu , dist_gini, dist_gni, dist_pub, dist_wrk_age, 
                        ncol = 2 , nrow = 3,
                        padding = unit(5,"line"))

```

```{r descriptive statistics , echo = FALSE , message = TRUE , warning = FALSE}
library(tableone)

hiv_table <-
tableone::CreateTableOne(data = pov_indicators_2018, 
                         vars = c("pc_poverty",
                                  "pc_higher_edu",
                                  "gni_per_capita",
                                  "gini_inequality",
                                  "pc_public_exp_of_gdp",
                                  "pc_working_age_pop"))

```

We can use QQ plots to establish how normal / non-normal the above plots actually are.

```{r}

# poverty
#dist_qq <-
pov_indicators_2018 %>%
  drop_na(pc_poverty) %>%
  ggplot(aes(sample = pc_poverty)) +
    stat_qq(aes(color = "Country"), size= 2, alpha = .5) + 
    scale_color_manual(values = "#1F456E" , name = "") +
    geom_abline(aes(intercept = mean(x = pc_poverty), 
                    slope     = sd(x = pc_poverty), 
                    linetype  = "Normally distributed"), color = "gray60", size = 1)  +
    scale_linetype_manual(values = "solid" , name = "") +
  labs(
    x = "Theoretical Normal Distribution",
    y = "Actual Poverty (%)"
  ) +
  theme(
    legend.position = "top",
    legend.justification = c(0,0)
  )

# gini
#dist_qq <-
pov_indicators_2018 %>%
  drop_na(gini_inequality) %>%
  ggplot(aes(sample = gini_inequality)) +
    stat_qq(aes(color = "Country"), size= 2, alpha = .5) + 
    scale_color_manual(values = "#1F456E" , name = "") +
    geom_abline(aes(intercept = mean(x = gini_inequality), 
                    slope     = sd(x = gini_inequality), 
                    linetype  = "Normally distributed"), color = "gray60", size = 1)  +
    scale_linetype_manual(values = "solid" , name = "") +
  labs(
    x = "Theoretical Normal Distribution",
    y = "Actual Gini (0-1)"
  ) +
  theme(
    legend.position = "top",
    legend.justification = c(0,0)
  )

# pun exp
#dist_qq <-
pov_indicators_2018 %>%
  drop_na(pc_public_exp_of_gdp) %>%
  ggplot(aes(sample = pc_public_exp_of_gdp)) +
    stat_qq(aes(color = "Country"), size= 2, alpha = .5) + 
    scale_color_manual(values = "#1F456E" , name = "") +
    geom_abline(aes(intercept = mean(x = pc_public_exp_of_gdp), 
                    slope     = sd(x = pc_public_exp_of_gdp), 
                    linetype  = "Normally distributed"), color = "gray60", size = 1)  +
    scale_linetype_manual(values = "solid" , name = "") +
  labs(
    x = "Theoretical Normal Distribution",
    y = "Actual Public Exp. (% GDP)"
  ) +
  theme(
    legend.position = "top",
    legend.justification = c(0,0)
  )

```

## Correlation
Using exploratory data analysis (EDA) and a simple linear model we can discover the trend between the y ~ x variables (here poverty ~ et. al).

```{r}
font_size <- 3
clr <- "#f68f46ff"

# theme
theme_reg <- function(){
      theme(
      legend.position = "none",
      legend.spacing.y = unit(-0.5, "mm"),
      panel.grid.minor = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 10),
      axis.line.x = element_line(color="black", size = 0.5),
      axis.line.y = element_line(color="black", size = 0.5),
      plot.margin = margin(0.5, 1, 0.5, 0.5, unit = "cm")
    )
}
  
```

At this point we've done a few things:

1. Used Histograms to visually look at kurtosis (skew)
2. Used Q plots to visually determine is the distribution is normal (an assumption of regression)
3. Used scatterplots to show the relation ship between our outcome (poverty rate) and explanatory variables 
4. Plotted a regression line (line of best fit)

Now we can detail the trends directly using correlation coefficients for all variables.

```{r detail trend , echo = FALSE , warning=FALSE}

pov_cor_2018 <- 
pov_indicators_2018 %>%
  drop_na() %>%
  summarise(pc_pov_vs_pc_he       = cor(pc_poverty, pc_higher_edu        ) ~round(),
            pc_pov_vs_gni_per_cap = cor(pc_poverty, gni_per_capita       ) ~round(),
            pc_pov_vs_gini        = cor(pc_poverty, gini_inequality      ) ~round(),
            pc_pov_vs_pc_pub      = cor(pc_poverty, pc_public_exp_of_gdp ) ~round(),  
            pc_pov_vs_pc_wrk_age  = cor(pc_poverty, pc_working_age_pop   ) ~round(),  
            sample_n              = n())

```

```{r}

# pov
corr_edu <-
pov_indicators_2018 %>%
  ggplot(aes(pc_higher_edu, pc_poverty)) +
    ggrepel::geom_text_repel(aes(size = "country", label = location), 
                             color = "#1F456E", alpha = 1, size = font_size, force = 1) +
    geom_smooth(aes(linetype = "Linear fit line") , method = "lm" , se = FALSE , color = clr , size = 0.5) +
    geom_text(label = paste0("cor: ", round(pov_cor_2018$pc_pov_vs_pc_he, digits = 4)), x = Inf, y = 0.025,
             hjust = 1, vjust = 0, size = font_size, color = clr) +
      guides(linetype = guide_legend(override.aes = list(size = 0.75 , alpha = 0.6))) +
      guides(size = guide_legend(override.aes = list(size = 6 , alpha = 0.5))) +
        scale_shape_manual (values = 17, name = "") + 
        scale_size_manual (values = 4, name = "") +
  scale_y_continuous(breaks = seq(0,0.2,0.05), limits = c(0,0.2), expand = c(0,0),
                     labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = seq(0,0.6,0.1), limits = c(0.15,0.6), expand = c(0,0),
                     labels = scales::percent_format(accuracy = 1)) +
    labs(x = "Population with Higher Education (%)", y = "Poverty Rate (%)") + 
  theme_reg() +
  theme(legend.position = c(0.9 , 0.9))


# gni
corr_gni <- 
pov_indicators_2018 %>%
  ggplot(aes(gni_per_capita, pc_poverty)) +
    ggrepel::geom_text_repel(aes(size = "country", label = location), 
                             color = "#1F456E", alpha = 1, size = font_size, force = 1) +
    geom_smooth(aes(linetype = "Linear fit line") , method = "lm" , se = FALSE , color = clr , size = 0.5) +
    geom_text(label = paste0("cor: ", round(pov_cor_2018$pc_pov_vs_gni_per_cap, digits = 4)), x = Inf, y = 0.025,
             hjust = 1, vjust = 0, size = font_size, color = clr) +
      guides(linetype = guide_legend(override.aes = list(size = 0.75 , alpha = 0.6))) +
      guides(size = guide_legend(override.aes = list(size = 6 , alpha = 0.5))) +
        scale_shape_manual (values = 17, name = "") + 
        scale_size_manual (values = 4, name = "") +
  scale_y_continuous(breaks = seq(0,0.2,0.05), limits = c(0,0.2), expand = c(0,0),
                     labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(expand = c(0,0), labels = scales::dollar_format()) +
    labs(x = "GNI / Capita (USD)", y = "Poverty Rate (%)") +
    theme_reg()

# gini
corr_gini <- 
pov_indicators_2018 %>%
  ggplot(aes(gini_inequality, pc_poverty)) +
    ggrepel::geom_text_repel(aes(size = "country", label = location), 
                             color = "#1F456E", alpha = 1, size = font_size, force = 1) +
    geom_smooth(aes(linetype = "Linear fit line") , method = "lm" , se = FALSE , color = clr , size = 0.5) +
    geom_text(label = paste0("cor: ", round(pov_cor_2018$pc_pov_vs_gini, digits = 4)), x = Inf, y = 0.025,
             hjust = 1, vjust = 0, size = font_size, color = clr) +
      guides(linetype = guide_legend(override.aes = list(size = 0.75 , alpha = 0.6))) +
      guides(size = guide_legend(override.aes = list(size = 6 , alpha = 0.5))) +
        scale_shape_manual (values = 17, name = "") + 
        scale_size_manual (values = 4, name = "") +
  scale_y_continuous(breaks = seq(0,0.2,0.05), limits = c(0,0.2), expand = c(0,0),
                     labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = seq(0,0.5,0.1), limits = c(0.2,0.5), expand = c(0,0)) +
    labs(x = "Gini Coefficient (0-1)", y = "Poverty Rate (%)") +
    theme_reg()

# public
corr_pub <-
pov_indicators_2018 %>%
  ggplot(aes(pc_public_exp_of_gdp, pc_poverty)) +
    ggrepel::geom_text_repel(aes(size = "country", label = location), 
                             color = "#1F456E", alpha = 1, size = font_size, force = 1) +
    geom_smooth(aes(linetype = "Linear fit line") , method = "lm" , se = FALSE , color = clr , size = 0.5) +
    geom_text(label = paste0("cor: ", round(pov_cor_2018$pc_pov_vs_pc_pub, digits = 4)), x = Inf, y = 0.025,
             hjust = 1, vjust = 0, size = font_size, color = clr) +
      guides(linetype = guide_legend(override.aes = list(size = 0.75 , alpha = 0.6))) +
      guides(size = guide_legend(override.aes = list(size = 6 , alpha = 0.5))) +
        scale_shape_manual (values = 17, name = "") + 
        scale_size_manual (values = 4, name = "") +
  scale_y_continuous(breaks = seq(0,0.2,0.05), limits = c(0,0.2), expand = c(0,0),
                     labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(expand = c(0,0), labels = scales::percent_format(accuracy = 1)) +
    labs(x = "Public Exp. (% GDP)", y = "Poverty Rate (%)") +
    theme_reg()

# working age
corr_wrk_age <- 
pov_indicators_2018 %>%
  ggplot(aes(pc_working_age_pop, pc_poverty)) +
    ggrepel::geom_text_repel(aes(size = "country", label = location), 
                             color = "#1F456E", alpha = 1, size = font_size, force = 1) +
    geom_smooth(aes(linetype = "Linear fit line") , method = "lm" , se = FALSE , color = clr , size = 0.5) +
    geom_text(label = paste0("cor: ", round(pov_cor_2018$pc_pov_vs_pc_wrk_age, digits = 4)), x = Inf, y = 0.025,
             hjust = 1, vjust = 0, size = font_size, color = clr) +
      guides(linetype = guide_legend(override.aes = list(size = 0.75 , alpha = 0.6))) +
      guides(size = guide_legend(override.aes = list(size = 6 , alpha = 0.5))) +
        scale_shape_manual (values = 17, name = "") + 
        scale_size_manual (values = 4, name = "") +
  scale_y_continuous(breaks = seq(0,0.2,0.05), limits = c(0,0.2), expand = c(0,0),
                     labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(expand = c(0,0), labels = scales::percent_format(accuracy = 1)) +
    labs(x = "Working Age Pop (% Total Pop)", y = "Poverty Rate (%)") +
    theme_reg()

# arrange
gridExtra::grid.arrange(corr_edu , corr_gini, corr_gni, corr_pub, corr_wrk_age, 
                        ncol = 2 , nrow = 3,
                        padding = unit(5,"line"))

```

We can see that the Gini coefficient is 0.8322 which is considered strong positive relationship (Zou et al., 2003).

## Modelling

```{r prediction workflow , echo = FALSE} 

# regression line of SSP distance per percentage point uninsured
mdl_pov_gini <-
lm(pc_poverty ~ gini_inequality , data = pov_indicators_2018 , na.action = na.exclude)

summary(mdl_pov_gini)

# confidence intervals (CI)
confint(mdl_pov_gini)

# prediction workflow
explantory_var <- expand_grid(gini_inequality = seq(0 , 1, 0.01))

prediction_data <-
  pov_indicators_2018 %>%
    mutate(
      prediction = predict(mdl_pov_gini , pov_indicators_2018),
      upper_ci   = predict(mdl_pov_gini , pov_indicators_2018 , interval = "confidence")[,2],
      lower_ci   = predict(mdl_pov_gini , pov_indicators_2018 , interval = "confidence")[,3]
    )

```

```{r}
# map prediction data onto sample data
residuals_mapped_lm <-
prediction_data %>%
  ggplot(aes(gini_inequality , pc_poverty, label = location)) +
    #geom_point(aes(size = "Country") , color = "#1F456E", alpha = 0.5) +
    scale_size_manual(values = 3, name = "") +
    geom_text(check_overlap = TRUE) +
    # manual calc of geom_smooth using lm() model
    geom_line(aes(x = gini_inequality , y = prediction , color = "Linear Fit Line") , size = 1 , alpha = 0.60) + 
    geom_smooth(color = "lightblue", linetype = "dashed", se = FALSE) +
    scale_color_manual(values = clr) +
    geom_segment(aes(xend = gini_inequality , yend = prediction , linetype = "Residual Error") , color = "grey75") +
    scale_y_continuous(breaks = seq(0,0.2,0.05), limits = c(0,0.2), expand = c(0,0),
                       labels = scales::percent_format(accuracy = 1)) +
    scale_x_continuous(breaks = seq(0,0.5,0.1), limits = c(0.2,0.5), expand = c(0,0)) +
    scale_linetype_manual(values = "dashed") +
    labs(x = "Gini Coefficient (0-1) [0 = pure equality, 1 = pure inequality]",
         y = "Poverty Rate (%)") +
  theme_reg()

residuals_mapped_lm

```

## NHST Step 1: Write the null and alternate hypotheses

H0: A model including the gini coefficient of a county is no better at explaining the poverty rate than a baseline model of the mean gini.

HA: A model including the gini coefficient in a county is better at explaining the poverty rate than a baseline model of the mean value of gini

## NHST Step 2: Compute the test statistic

The test statistic for this model is F and its value is F(1, 32) = 59.41.

## NHST Step 3: Calculate the probability that your test statistic is at least as big as it is if there is no relationship (i.e., the null is true) 

There is a tiny probability (p < .001) of an F as big as 59.41 or bigger if the null hypothesis were true.

## NHST Steps 4 and 5: Interpret the probability and write a conclusion

Given the tiny p-value, we reject the null hypothesis in favor of the alternate hypothesis that gini is helpful in explaining the poverty rate of a county.




# Assumptions

The calculations underlying the simple linear regression model of Pc_pov ~ gini are based on several assumptions about the data used in the model: 

* Observations are independent.
* The outcome is continuous.
* The relationship between the two variables is linear (linearity). 


* The variance is constant with the points distributed equally around the line (homoscedasticity). 

lmtest::bptest(formula = pov_indicators_2018$pc_poverty ~ pov_indicators_2018$gini_inequality)

BP = 0.84642, df = 1, p-value = 0.3576

ASSUMPTION MET: The Breusch-Pagan test statistic has a large p-value (BP = 0.85; p < .5), indicating that the null hypothesis of constant variance would be accepted.

* The residuals are independent. 

lmtest::dwtest(formula = pov_indicators_2018$pc_poverty ~ pov_indicators_2018$gini_inequality)

DW = 1.8166, p-value = 0.2667

ASSUMPTION MET: The D-W statistic was near 2 and the p-value was high, so we concluded that the null hypothesis is retained.

* The residuals are normally distributed.

ASSUMPTION MET: The Q-Q plot suggests the residuals are the same from the values you’d expect from a normal distribution.



***

R markdown allows to easily format your text. You can add [links](www.r-graph-gallery.com), write in **bold** or *italic*.

Further formatting can be found [here](https://holtzy.github.io/Pimp-my-rmd/)

***