---
title: "UK Deprivation Table"
author: "NearAndDistant"
date: '2022-06-06'
output: html_document
---

```{r}
dir.create(here::here("uk_deprivation_index"))

# temp file to be used throughout
temp   <- tempfile()
```

# DEPRIVATION DATA

Data from MySociety, here: https://github.com/mysociety/composite_uk_imd
Methodology from, here: https://bmjopen.bmj.com/content/bmjopen/6/11/e012750.full.pdf

The composite labels vary by national dataset, but the key columns in the UK England dataset are:

lsoa                  - The code for the LSOA/Datazone/SOA
original_decile       - the original decile in the four different IMD datasets.
overall_local_score   - the original deprivation score in the four different IMD datasets.
UK_IMD_E_score        - the transformed deprivation score into the combined index. This will be the same as the original for the base country.
E_expanded_decile     - the English deciles, with scores for other nations fitted into the same devisions as the original deciles.
UK_IMD_E_pop_decile   - A new set of deciles for the whole dataset, so that 10% of the population is in each decile (uneven number of areas).
UK_IMD_E_pop_quintile - A new set of quintiles for the whole dataset, so that 20% of the population is in each quintile (uneven number of areas).

The ONS do a subset of this analysis for England and Wales **only**. We want all UK nations (inc. Scotland & N. Ireland), see here:
https://commonslibrary.parliament.uk/research-briefings/cbp-7327/

```{r}
library(tidyverse)

ukdi_raw <- 
  read_csv("https://raw.githubusercontent.com/mysociety/composite_uk_imd/master/uk_index/UK_IMD_E.csv") %>% 
  janitor::clean_names() %>%
  select(nation, lsoa, uk_imd_e_rank, uk_imd_e_score)

```

# CONSTITUENCY POPULATION

### Constituencies (LSOA):

533 in England (32,844)
40  in Wales (1,909)
59  in Scotland (6,976)
18  in Northern Ireland (890)

Here we will be using the joined dataset, for this project and filtering to what we need for all 650 constituencies, here:
https://commonslibrary.parliament.uk/constituency-statistics-population-by-age/

# POPULATION ESTIMATES

```{r}
# Constituencies population

onsurl <- "https://data.parliament.uk/resources/constituencystatistics/PowerBIData/Demography/Population.xlsx"
temp   <- curl::curl_download(url = onsurl, destfile = temp, quiet = FALSE, mode = "wb")

con_pop_raw <- 
  readxl::read_xlsx(temp, sheet = "Age by year data" , range = "A1:E59151") %>% 
  janitor::clean_names() %>%
  distinct(pcon11cd, .keep_all = TRUE) %>%
  select(con_code = "pcon11cd", con_name = "pcon11nm", pop = "all_ages") %>%
  mutate(weight = pop / sum(pop))

```

```{r}
# SLOA population

# England and Wales (34,753 = does not match IMD file)
# source: https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/lowersuperoutputareamidyearpopulationestimatesnationalstatistics
curl <- "https://www.ons.gov.uk/file?uri=%2fpeoplepopulationandcommunity%2fpopulationandmigration%2fpopulationestimates%2fdatasets%2flowersuperoutputareamidyearpopulationestimates%2fmid2020sape23dt2/sape23dt2mid2020lsoasyoaestimatesunformatted.xlsx"
temp   <- curl::curl_download(url = curl, destfile = temp, quiet = FALSE, mode = "wb")

lsoa_pop_eng <- 
  readxl::read_xlsx(temp, sheet = "Mid-2020 Persons", skip = 4) %>% 
  janitor::clean_names() %>%
  distinct(lsoa_code, .keep_all = TRUE) %>%
  select(lsoa_code, pop_est = "all_ages")

# Scotland (6,976 = the IMD file)
# source: https://www.opendata.nhs.scot/dataset/population-estimates/resource/c505f490-c201-44bd-abd1-1bd7a64285ee?inner_span=True
lsoa_pop_scot <- 
  read_csv("uk_deprivation_index/c505f490-c201-44bd-abd1-1bd7a64285ee.csv") %>%
  janitor::clean_names() %>%
  filter(year == 2020) %>%
  distinct(data_zone, .keep_all = TRUE) %>%
  select(lsoa_code = "data_zone", pop_est = "all_ages")

# N. Ireland (890 = the IMD file)
# source: https://data.gov.uk/dataset/2135b9bf-eb38-4c81-b614-7b07e8fdfc82/population-estimates-for-super-output-areas-soas-and-former-electoral-wards-northern-ireland/datafile/fc3f91dd-a0c7-488a-9996-bd057a7cfffe/preview
lsoa_pop_nire <-
  read_csv("https://www.opendatani.gov.uk/dataset/2135b9bf-eb38-4c81-b614-7b07e8fdfc82/resource/fc3f91dd-a0c7-488a-9996-bd057a7cfffe/download/super-output-areas-soas-by-gender-and-broad-age-bands-mid-2001-to-mid-2020.csv") %>%
  janitor::clean_names() %>%
  filter(mid_year_ending == 2020 & gender == "All persons" & age_group == "All ages") %>%
  distinct(geo_code, .keep_all = TRUE) %>%
  select(lsoa_code = "geo_code", pop_est = "population_estimate")

```

```{r}
# bind all SLOA dfs together
lsoa_pop_all <- rbind(lsoa_pop_eng, lsoa_pop_scot, lsoa_pop_nire)
```

# MAPPING LSOA TO CONSTITUENCIES

We need to join the deprivation index data - which is in Lower Super Output Areas (LSOA) - to their respective constituencies.

Scottish data, here: https://statistics.gov.scot/data/data-zone-lookup
N. Ireland data, here: https://www.nisra.gov.uk/support/geography/northern-ireland-small-areas
England & Wales data, here: https://geoportal.statistics.gov.uk/search?collection=Dataset&sort=name&tags=all(LUP_WPC) 
**English & Wales data is not 1:1 therefore centroid methodology used**

### LSOAs:

32,844  in England
 1,909  in Wales
 6,976  in Scotland
   890  in Northern Ireland.
   
42,619  Total LSOAs

```{r}

# Scotland
lsoa_scot_raw <- read_csv("https://scottish-government-files.s3.amazonaws.com/13360f3a-ca68-4f3e-8b7f-caffed8712eb/DataZone2011lookup_2022-05-31.csv")

# ukdi Scotland
lsoa_scot <- 
ukdi_raw %>%
  filter(nation == "S") %>%
  left_join(lsoa_scot_raw, by = c(lsoa = "DZ2011_Code")) %>%
  select(nation, lsoa, lsoa_nm = IZ2011_Name, con_code = UKPC_Code, con_name = UKPC_Name, uk_imd_e_rank, uk_imd_e_score)

```

```{r}

# N. Ireland
## missing names (codes only, we need names for graphics) so download seperate table to join

# lsoa codes
nisurl <- "https://www.nisra.gov.uk/sites/nisra.gov.uk/files/publications/Look-up%20Tables_0.xlsx"
temp   <- curl::curl_download(url = nisurl, destfile = temp, quiet = FALSE, mode = "wb")

lsoa_nire_cd <- readxl::read_xlsx(temp, sheet = "COA2001") %>% 
  janitor::clean_names() %>%
  distinct(soa2001, .keep_all = TRUE) %>%
  select(soa2001, aa2008)

# lsoa names, source, here: https://data.nicva.org/dataset/geographical-names-and-codes-tables/resource/f072a627-5e70-4647-8b55-e5cdc1ba6b9a
lsoa_nire_nm <- read_csv("https://data.nicva.org/sites/default/files/ni_geography_names_codes_all_types_Mar2017.csv")

# join nire codes to names
lsoa_nire_cd_nm <-
  lsoa_nire_cd %>%
  left_join(lsoa_nire_nm, by = c(soa2001 = "code")) %>%
  rename(soa2001nm = name) %>%
  left_join(lsoa_nire_nm, by = c(aa2008 = "code")) %>%
  rename(aa2008nm = name) %>%
  select(-unit.x, -unit.y, -source.x, -source.y)

# ukdi N. IRE
lsoa_nire <- 
  ukdi_raw %>%
  filter(nation == "N") %>%
  left_join(lsoa_nire_cd_nm, by = c(lsoa = "soa2001")) %>%
  select(nation, lsoa, lsoa_nm = soa2001nm, con_code = aa2008, con_name = aa2008nm, uk_imd_e_rank, uk_imd_e_score)

```

### ENGLAND LSOAs to Constituencies

For joining LSOAs to Constituencies there is not a 1:1 relationship within *England only* i.e. a LSOA is 1:1 in Wards but a Ward != 1:1 with Constituencies. We therefore need to use a geo-join to establish which LSOAs are within which Constituency using a fuzzy methodology.

Alasdair Rae has created a single shapefile that covers the different sets of small area that will be helpful in cross-border mapping.
Source: https://drive.google.com/drive/folders/1L-C-Ym7VTwRhIRDQUMHPxFvOznMrKfy3

Joining points intersections of polygons:
https://ryanpeek.org/2019-04-29-spatial-joins-in-r/

```{r}
library(sf)

# constituency areas
cons_shp <- readRDS(url("https://www.dropbox.com/s/XXX/uk_cons_shp?dl=1")) # personal dropbox

# convert to simple features (sf)
cons_sf <- 
  cons_shp %>% 
  st_as_sf() %>% 
  select(pcon19cd, pcon19nm)


# lsoa area centroids (32,844)

# December 2011 is latest data held by ONS - not perfect given our lsoa data is from 2020.
# source: https://geoportal.statistics.gov.uk/maps/b7c49538f0464f748dd7137247bbc41c/about
lsoa_cent_raw   <- read_csv(here::here("uk_deprivation_index/Lower_Layer_Super_Output_Areas_(December_2011)_Population_Weighted_Centroids.csv"))

lsoa_ukdi_centr <- 
  ukdi_raw %>% 
  filter(nation %in% c("E", "W")) %>% 
  left_join(lsoa_cent_raw, by = c(lsoa = "lsoa11cd")) %>% 
  st_as_sf(coords = c("X", "Y"), crs = st_crs(cons_sf))

# Overlay points and extract just the code column: 
lsoa_geoj <- lsoa_ukdi_centr %>% st_join(cons_sf, left = TRUE, largest = TRUE)

```

```{r}

# England - lsoa's and constituency geo-areas are removed by degrees from each other. We use Wards to map them across to each other.

## constituencies to wards
con_ward  <- read_csv(here::here("uk_deprivation_index/Ward_to_Westminster_Parliamentary_Constituency_to_Local_Authority_District_to_Upper_Tier_Local_Authority__December_2020__Lookup_in_the_United_Kingdom_V2.csv"))

## wards to lsoa
ward_lsoa <- read_csv(here::here("uk_deprivation_index/Lower_Layer_Super_Output_Area__2011__to_Ward__2020__Lookup_in_England_and_Wales.csv"))

# lsoa_eng  <- 
#   ward_lsoa %>% 
#   left_join(con_ward, by = c("WD20CD", "LAD20CD")) %>% 
#   mutate(LSOA11NM = str_sub(LSOA11NM, start = 0, end = str_length(LSOA11NM)-4)) %>%
#   distinct(LSOA11CD, .keep_all = TRUE) %>%
#   left_join(con_ward2, by = c(WD20CD = "WD19CD")) %>%
#   # there have been some changes and some fallouts which we need, we will keep the latest changes and add fallouts
#   mutate(check = PCON19CD == PCON20CD) %>%
#   mutate(PCON20CD = if_else(is.na(PCON20CD), PCON19CD, PCON20CD), 
#          PCON20NM = if_else(is.na(PCON20NM), PCON19NM, PCON20NM)) %>% 
#   select(LSOA11CD, LSOA11NM, PCON20CD, PCON20NM)

lsoa_geoj_miss <- 
lsoa_geoj %>%
  filter(is.na(pcon19cd)) %>%
  left_join(ward_lsoa, by = c(lsoa = "LSOA11CD")) %>% 
  left_join(con_ward, by = c("WD20CD", "LAD20CD")) %>%
  mutate(pcon19cd = if_else(is.na(pcon19cd), PCON20CD, pcon19cd),
         pcon19nm = if_else(is.na(pcon19nm), PCON20NM, pcon19nm)) %>%
  select(names(lsoa_geoj))

# final for England & Wales
lsoa_eng <-
lsoa_geoj %>%
  filter(!is.na(pcon19cd)) %>%
  mutate(lsoa11nm = str_sub(lsoa11nm, start = 0, end = str_length(lsoa11nm)-4)) %>%
  mutate(pcon19nm = if_else(pcon19nm == "Ynys Mon", "Ynys Môn", pcon19nm)) %>%
  rbind(lsoa_geoj_miss) %>%
  select(-objectid, nation, lsoa, lsoa_nm = lsoa11nm, con_code = pcon19cd, con_name = pcon19nm, uk_imd_e_rank, uk_imd_e_score) %>%
  st_drop_geometry()

```

```{r}
# joins lsoas

ukdi_cons <- 
  rbind(lsoa_eng, lsoa_scot, lsoa_nire) %>%
  # add sloa pop
  left_join(lsoa_pop_all, by = c(lsoa = "lsoa_code")) %>%
  # add constituency pop for weights
  left_join(con_pop_raw) %>%
  select(names(lsoa_eng), lsoa_pop = pop_est, con_pop = pop)

```

```{r}
# checks no missing values
ukdi_cons %>% anti_join(con_pop_raw, by = "con_code") %>% count(con_name)
# checks by how much lsoa pop is from con pop
ukdi_cons %>% group_by(con_code) %>% mutate(lsoa_con_pop = sum(lsoa_pop), check = abs(con_pop - lsoa_con_pop)) %>% view()
```

```{r}
# creates weights and reindexes for constituencies

ukdi_index <-
ukdi_cons %>%
  group_by(con_code, con_name) %>%
  mutate(lsoa_weight = lsoa_pop / con_pop,
         ukdi_weight = uk_imd_e_score * lsoa_weight) %>%
  summarise(uk_imd_con_score = sum(ukdi_weight)) %>%
  ungroup() %>%
  arrange(desc(uk_imd_con_score)) %>%
  group_by(fct_inorder(con_code)) %>%
  mutate(cons_rank     = cur_group_id()) %>%
  ungroup()

```

```{r}
# checks all 650 constituencies caught
con_pop_raw %>% left_join(ukdi_index, by = "con_code") %>% count() %>% mutate(check = n == 650)
```

# ELECTION DATA

```{r}

# Constituency election data, here: https://www.theyworkforyou.com/mps/
curl <- "https://www.theyworkforyou.com/mps/?f=csv"
temp   <- curl::curl_download(url = curl, destfile = temp, quiet = FALSE, mode = "wb")

election21_raw <- read_csv(temp) %>% janitor::clean_names()

# checks we can join by constituency name as the MySociety data does not have the codes
election21_raw %>% anti_join(ukdi_index_summary, by = c(constituency = "PCON20NM"))

# Weston-Super-Mare is the only disjoin due to "Super" and "super" spelling
election21_cln <- 
  election21_raw %>% 
  mutate(constituency = if_else(constituency == "Weston-super-Mare", "Weston-Super-Mare", constituency)) %>%
  mutate(abb = case_when(
  party == "Conservative"                       ~ "CON"   ,
  party == "Labour"                             ~ "LAB"   ,
  party == "Labour/Co-operative"                ~ "LAB"   ,    # Labor/Co-Op normally included in Labor totals
  party == "Scottish National Party"            ~ "SNP"   ,
  party == "Liberal Democrat"                   ~ "LD"   ,
  party == "DUP"                                ~ "DUP"   , 
  party == "Sinn Féin"                          ~ "SF"    , 
  party == "Plaid Cymru"                        ~ "PLAID" ,
  party == "Social Democratic and Labour Party" ~ "SDLP"  , 
  party == "Alba"                               ~ "ALB"   ,
  party == "Independent"                        ~ "IND"   , 
  party == "Alliance"                           ~ "A"     , 
  party == "Green"                              ~ "GRN"   ,
  party == "Speaker"                            ~ "ORDER!")) %>%
  mutate(fill = case_when(
  party == "Conservative"                       ~ "#0087DC",
  party == "Labour"                             ~ "#E4003B",
  party == "Labour/Co-operative"                ~ "#E4003B",    # Labor/Co-Op normally included in Labor totals
  party == "Scottish National Party"            ~ "#fff481",
  party == "Liberal Democrat"                   ~ "#FDBB30",
  party == "DUP"                                ~ "#a3033b", 
  party == "Sinn Féin"                          ~ "#258c45", 
  party == "Plaid Cymru"                        ~ "#8ab47b",
  party == "Social Democratic and Labour Party" ~ "#1b3a19", 
  party == "Alba"                               ~ "#005db8",
  party == "Independent"                        ~ "#808080", 
  party == "Alliance"                           ~ "#FDD017", 
  party == "Green"                              ~ "#78B943",
  party == "Speaker"                            ~ "#363151"))

# there are two other constituencies (Tiverton and Wakefield) without data due to resignations. Their by-elections are due 23 June (at time of writing). 
# See here: https://www.parliament.uk/about/how/elections-and-voting/by-elections/by-elections-2019/

# order of parties
order  <- election21_cln %>% count(abb, fill)  %>% arrange(desc(n)) %>% mutate(abb  = factor(abb,  ordered = TRUE),
                                                                               fill = factor(fill, ordered = TRUE))
```

```{r}
# joins constituency and election data

ukdi_cons_mps <-
  ukdi_index %>%
  left_join(election21_cln, by = c(con_name = "constituency")) %>%
  mutate(con_name = if_else(con_name == "East Kilbride, Strathaven and Lesmahagow", "East Kilbride & Strathaven", con_name),
         con_name = if_else(con_name == "Runnymede and Weybridge", "Runnymede", con_name),
         con_name = if_else(con_name == "Dumfriesshire, Clydesdale and Tweeddale", "Dumfriesshire & Clydesdale", con_name)) %>%
  mutate(abb  = factor(abb,  levels = order$abb, ordered = TRUE),
         fill = factor(fill, levels = order$fill, ordered = TRUE))

```

# CREATING THE PLOT

The map is essentially a giant tile map, filled by party.

```{r}
# we want to record the making of this plot so this is a setup
library(camcorder)

gg_record(
  dir    = here::here("uk_deprivation_index/recording"), # where to save the recording
  device = "png", # device to use to save images
  width  = 10, # width of saved image
  height = 16, # height of saved image
  units  = "in", # units for width and height
  dpi    = 300 # dpi to use when saving image
)

# to chnage any of the formats going forward use the below
gg_resize_film(
  height = 16,
  width  = 11.5,
  units  = "in",
  dpi    = 360
)

```

```{r}
# text

what  <- "There are 650 constituencies in the UK, which we rank by deprivation from top left (most deprived) to bottom right (least deprived)."
what2 <- "The first column contains the 65 most deprived constituencies, the second one has the next 65 most deprived, and so on. The numbers along the top indicate which decile each constituency is in. The UK-wide deprivation data used is published by Alex Parsons (@alexparsons) of mySociety, following earlier work by Abel et al. (2016). It uses the most recent deprivation indices for all parts of the UK and there are a number of different datasets available."
what3 <- "The original graphic was created by Alasdair Rae (@undertheradar) and any errors here are strictly my own."

note <- "This graphic is up date as of 3 March 2022, following the result of the Birmingham, Erdington by-election."

meth  <- "Deprivation scores are taken from mySociety data which uses lower super output areas (lsoa)."
meth2 <- "The constituency deprivation ranks are created by aggregating lsoa by constituency. For Scotland and N. Ireland this is complete through their respective lsoa lookup tables as they have a 1:1 relationship to constituencies. For England and Wales lsoa do not have a 1:1 relationship to constituencies therefore (following Alasdair Rae's method) ranks were created using population weighted centroids to establish the constituency 'best-fit'."
meth3 <- "Once grouped by constituency deprivation scores were weighted by their population contribution to that constituency to give a constituency score which we then rank."
meth4 <- "There is a great deal of variation within each constituency but those showing as most deprived have the UKs largest share of deprived neighbourhoods. From a methodological point of view the exact positioning of each constituency would change dependent on which dataset / method chosen (this is evidenced by the differences shown between Alasdair's and my own graphic) but the decile positioning is mostly the same."

source1 <- "Deprivation data: githubusercontent.com/mysociety/composite_uk_imd/master/uk_index/UK_IMD_E"
source2 <- "Pop weighted centroids (england): data.gov.uk/dataset/a40f54f7-b123-4185-952f-da90c56b0564/"
source3 <- "Election reults: theyworkforyou.com/mps/"
  
```


```{r}
library(ggh4x) # https://teunbrand.github.io/ggh4x/articles/Miscellaneous.html#string-legends-1
# font_add_google("Special Elite", "type")

plot_ukdi <- 
ukdi_cons_mps %>%
  # formatting
  mutate(color = if_else(party == "Scottish National Party", "grey", "white")) %>% # snp yellow has little contrast with white so change text color
  mutate(color = if_else(is.na(party), "grey",  color)) %>% # any cons without MPs (at by-election)
  mutate(con_name = str_replace(con_name, " and ", " & ")) %>% # reduce long text
  mutate(x_pos  = rep(1:10, each = 65),
         y_pos  = rep(1:65, time = 10)) %>%
  # plot
  ggplot(aes(x_pos, y_pos, fill = fill, label = str_wrap(con_name,21))) + # control wrapping strings to fit into tile
  geom_tile(color = "white", width = 0.95, height = 0.98, alpha = 0.95) +
  geom_text(aes(color = color), size = 1.50, hjust = 0, nudge_x = -0.455, show.legend = FALSE) + 
  scale_fill_identity(NULL, guide = "stringlegend", labels = levels(ukdi_cons_mps$abb)) +
  scale_color_manual(values = c("#A7AFB2", "white"), na.value = "white") +
  guides(fill  = guide_stringlegend(nrow = 1, face = "bold")) +
  scale_x_continuous(breaks = c(1:10), expand = c(0.01,0.01), position = "top") +
  scale_y_reverse(breaks = c(1:65), expand = c(0.01,0.01), sec.axis = sec_axis(~., breaks = c(1:65), labels = c(586:650))) +
  coord_cartesian(clip = "off", ylim = c(65, 0)) +
  labs(title   = "UK CONSTITUENCIES BY\nDEPRIVATION, 2022", 
       caption = "@NearAndDistant") +
  annotate("text", label = "MORE DEPRIVED",  x =  1, y = 66.5, hjust = 0.3, family = "Montserrat Semibold", color = "#696969") +
  annotate("text", label = "LEAST DEPRIVED", x = 10, y = 66.5, hjust = 0.7, family = "Montserrat Semibold", color = "#696969") +
  annotate("text", label = "DATA SOURCES",   x =  1, y = 71, hjust = 0.30, family = "Montserrat Semibold", color = "#696969") +
  annotate("text", label = source1, x =  1, y = 72, hjust = 0.06,   family = "Montserrat Semibold", color = "#bebebe", size = 3) +
  annotate("text", label = source2, x =  1, y = 73, hjust = 0.0625, family = "Montserrat Semibold", color = "#bebebe", size = 3) +
  annotate("text", label = source3, x =  1, y = 74, hjust = 0.145,  family = "Montserrat Semibold", color = "#bebebe", size = 3) +
  theme_void() +
  theme(plot.title   = element_text(size = 40, hjust = 0.5,  vjust =   3, family = "Montserrat Semibold", color = "#454545"),
        plot.caption = element_text(size = 15, hjust = 1.345, vjust = -32, family = "Montserrat Bold",     color = "#a7a7a7"),
        legend.position  = c(0.525, -0.045),
        legend.direction = "horizontal",
        legend.key       = element_blank(),
        legend.text      = element_text(size = 16, family = "Montserrat"),
        axis.text.x       = element_text(size = 8, family = "Montserrat Semibold", color = "#e5eaed", vjust = -6),
        axis.text.y.left  = element_text(size = 6, family = "Montserrat Semibold", color = "#e5eaed"),
        axis.text.y.right = element_text(size = 6, family = "Montserrat Semibold", color = "#e5eaed"),
        plot.margin = margin(2, 5, 4.5, 5, unit = "cm"),
        panel.background = element_rect(fill = "white", color = "white"),
        plot.background  = element_rect(fill = "white", color = "#d7d7d7", size = 8))

ggsave(plot = plot_ukdi, here::here("uk_deprivation_index/graphic_no_ann.png"), width = 11.5, height = 16, dpi = 360)

```

```{r}
library(cowplot)

plot_final <-
ggdraw() +
  draw_image(here::here("uk_deprivation_index/graphic_no_ann.png")) +
  # explainer
  annotate("text", label = "WHAT IS THIS?",     x = 0.16, y = 0.85,  hjust = 1, family = "Montserrat Semibold", color = "#696969") +
  annotate("text", label = str_wrap(what , 24), x = 0.16, y = 0.805, hjust = 1, family = "Montserrat Semibold", color = "#bebebe", size = 3) +
  annotate("text", label = str_wrap(what2, 24), x = 0.16, y = 0.61,  hjust = 1, family = "Montserrat Semibold", color = "#bebebe", size = 3) +
  annotate("text", label = str_wrap(what3, 24), x = 0.16, y = 0.45,  hjust = 1, family = "Montserrat Semibold", color = "#bebebe", size = 3) +
  # note
  annotate("text", label = "NOTE",              x = 0.84, y = 0.85,  hjust = 0, family = "Montserrat Semibold", color = "#696969") +
  annotate("text", label = str_wrap(note , 24), x = 0.84, y = 0.815, hjust = 0, family = "Montserrat Semibold", color = "#bebebe", size = 3) +
  # methodology
  annotate("text", label = "METHODOLOGY",        x = 0.84, y = 0.725, hjust = 0, family = "Montserrat Semibold", color = "#696969") +
  annotate("text", label = str_wrap(meth ,  24), x = 0.84, y = 0.69, hjust = 0, family = "Montserrat Semibold", color = "#bebebe", size = 3) +
  annotate("text", label = str_wrap(meth2 , 24), x = 0.84, y = 0.53, hjust = 0, family = "Montserrat Semibold", color = "#bebebe", size = 3) +
  annotate("text", label = str_wrap(meth3 , 24), x = 0.84, y = 0.35, hjust = 0, family = "Montserrat Semibold", color = "#bebebe", size = 3) +
  annotate("text", label = str_wrap(meth4 , 24), x = 0.84, y = 0.19, hjust = 0, family = "Montserrat Semibold", color = "#bebebe", size = 3)

ggsave(plot = plot_final, here::here("uk_deprivation_index/final_graphic.png"), width = 11.5, height = 16, dpi = 360)

```
