---
title: "Income & Benefits Distributions"
author: "NearAndDistant"
date: "25/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Grab ONS Gross Data from GitHub (Script in Repository)

Gross Disposable household Income is the amount of money that that all of the individuals in the household sector have available for spending or saving after income distribution measures (for example, taxes, social contributions and benefits) have taken effect. GDHI does not provide measures relating to actual households or family units. The figures cover regions, sub-regions and local areas of the UK. It consists of:

##### Primary resources
0: Operating surplus
1: Mixed income
2: Compensation of employees
3: Property income, received

= Primary Resource total

#### Importing
```{r}
library(tidyverse)

# you can see the primary script to grab this at https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gross_income_1997_to_2018/
ons_gdhi_components_perhead_raw <-
read_csv("https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gdhi_97_18/ons_gdhi_perhead_components_97_18.csv")

gdhi_comp_indi_raw <-
read_csv("https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gdhi_97_18/gdhi_perhead_components_indices.csv")


```

#### Palette
```{r}

# create tibble for colours
nuts1_palette <- tibble(region_name = c("Scotland", "North East", "North West", "Yorkshire", "East Midlands", "West Midlands",
                                        "Wales", "South West", "South East", "London", "East of England", "Northern Ireland"),
                        fill        = c("#9E7BB5", "#B7E4F9FF", "#24325FFF", "#82491EFF", "#FB6467FF", "#FF6348FF", 
                                        "#ADE2D0FF", "#E89242FF", "#FAE48BFF", "#3F4041FF", "#917C5DFF", "#526E2DFF"))

library(ggsci)

region_palette <- c("#9E7BB5", "#B7E4F9FF" , "#24325FFF", "#82491EFF", "#FB6467FF", "#FF6348FF",
                    "#ADE2D0FF", "#E89242FF", "#FAE48BFF", "#3F4041FF", "#917C5DFF", "#526E2DFF")

scales::show_col(Redmonder::redmonder.pal(9, "sPBIBu"))

palette       <- Redmonder::redmonder.pal(5, "sPBIBu")

library(rcartocolor)

scales::show_col(rev(carto_pal(7, "TealRose")))

palette       <- rev(carto_pal(7, "TealRose"))

```

##### Cleaning & Wrangle
```{r}

region_order <- c("Scotland", "North East", "North West", "Yorkshire", "East Midlands", "West Midlands",
                  "Wales", "South West", "South East", "London", "East of England", "Northern Ireland")

gdhi_income <- c("Operating surplus", "Mixed income", "Compensation of employees", "Property income, received", "Primary resources total")

```

###########
Table
###########

```{r}

gdhi_reg_inc_indi_cln <-
gdhi_comp_indi_raw %>%
  janitor::clean_names(numerals = "left") %>%
  filter(nuts_level %in% c("NUTS1", "UK")) %>% # NUTS1 level is regions
  pivot_longer(cols = starts_with("x") , names_to = "year" , values_to = "value") %>%
  mutate(year = str_remove(year , "x"),
         year = as.numeric(year)) %>%
  mutate(region_name = if_else(region_name == "Yorkshire and The Humber", "Yorkshire", region_name)) %>%
  select(-nuts_level, -nuts_code, -transaction_code) %>%
  left_join(nuts1_palette) %>%
  mutate(region_name = as.factor(region_name),
         region_name = factor(region_name, levels = c("United Kingdom", region_order)))


gdhi_reg_inc_indi <-
  gdhi_reg_inc_indi_cln %>%
  filter(transaction %in% gdhi_income) # filter all unnecessary transactions for this analysis

```

```{r}
library(gt)
library(gtExtras)

table_income_indices <-
gdhi_reg_inc_indi %>%
  filter(transaction == "Primary resources total") %>%
  arrange(region_name) %>%
  group_by(region_name) %>%
  mutate(trend = list(value)) %>%
  ungroup() %>%
    pivot_wider(id_cols = c(region_name, trend), names_from = year, values_from = value) %>%
      gt(rowname_col = "region_name") %>%
      gt_sparkline(trend, width = 25) %>%
      grand_summary_rows(columns = c(`1997`:`2018`), fns = list(Range = ~max(.) - min(.)), use_seps = TRUE, missing_text = "") %>%
      #### color scale
      data_color(columns = c(`1997`:`2018`),
                 # custom defined values - notice that order matters!
                 colors = scales::col_factor(palette = palette, domain = NULL), 
                 alpha = 0.4, autocolor_text = FALSE) %>% 
      tab_style(style = list(cell_fill(color = "white")), locations = cells_body(rows = 1)) %>%
      #### cell boarder format
      tab_style(style = list(cell_borders(sides = "right", color = "black", weight = px(3))),
                locations = list(cells_body(columns = trend))) %>%
      tab_style(style = list(cell_borders(sides = "top", color = "grey60", weight = px(1))),
                locations = list(cells_column_labels(columns = everything()))) %>%
      tab_style(style = list(cell_borders(sides = "bottom", color = "black", weight = px(3))),
                locations = list(cells_column_labels(columns = everything()))) %>%
      tab_source_note("Data: ONS | Table: @NearAndDistant") %>%
      tab_header(title = md("**Monthly Gross Income from 1997 to 2018 as Indices of UK Average; UK Average = 100**")) %>%
      gt_theme_nytimes() 

table_income_indices

```

```{r}

# Writing to File
gtsave_extra(data = table_income_indices, filename = here::here("projects/ons_region_gdhi_1997_2018/outputs/gdhi_income_table.png"), 
             zoom = 1.5 , vwidth = 1400, vheight = 500)

```

#################################
Income Analysis
#################################

```{r}

gdhi_regional_inc_cln <-
ons_gdhi_components_perhead_raw %>%
  janitor::clean_names(numerals = "left") %>%
  filter(nuts_level %in% c("NUTS1" , "UK")) %>% # NUTS1 level is regions
  pivot_longer(cols = starts_with("x") , names_to = "year" , values_to = "value") %>%
  mutate(year = str_remove(year , "x"),
         year = as.numeric(year)) %>%
  mutate(region_name = if_else(region_name == "Yorkshire and The Humber", "Yorkshire", region_name)) %>%
  select(-nuts_level, -nuts_code, -transaction_code) %>%
  mutate(monthly_value = round(value / 12, 1)) %>%
  # rate of change summary for sec_axis_labels 
  group_by(region_name, transaction) %>%
  mutate(region_ror = (last(monthly_value) - first(monthly_value)) / first(monthly_value)) %>%
  mutate(sec_axis_labels = paste0(region_name, ", ", scales::comma(monthly_value, prefix = "£", accuracy = 1) , " (+", scales::percent(region_ror, accuracy = 2),")")) %>%
  ungroup() %>%
  left_join(nuts1_palette) %>%
  mutate(region_name = as.factor(region_name),
         region_name = factor(region_name, levels = c("United Kingdom" , region_order)))

gdhi_regional_inc <-
  gdhi_regional_inc_cln %>%
    filter(transaction %in% gdhi_income) # filter all unnecessary transactions for this analysis

```

```{r}
theme_set(theme_minimal())

#### secondary axis annotations
east_midlands_coord <-
gdhi_regional_inc$monthly_value[which(gdhi_regional_inc$year == 2018 & gdhi_regional_inc$region_name =="East Midlands" & gdhi_regional_inc$transaction == "Primary resources total")]

west_midlands_coord <-
gdhi_regional_inc$monthly_value[which(gdhi_regional_inc$year == 2018 & gdhi_regional_inc$region_name =="West Midlands"  & gdhi_regional_inc$transaction == "Primary resources total")]
####

plot_region_inc_timeseries <-
gdhi_regional_inc %>%
  filter(region_name != "United Kingdom") %>%
  filter(transaction == "Primary resources total") %>%
  ggplot() +
  geom_line(aes(year , monthly_value , color = fill, group = region_name), show.legend = TRUE , size = 1.5) +
  scale_x_continuous(breaks = c(1997,seq(2000,2015,5),2018), limits = c(1997,2018), expand = c(0,0)) +
  expand_limits(y = 0) +
  scale_y_continuous(labels = scales::comma_format(accuracy = 1 , prefix = "£"), expand = c(0,0),
                     sec.axis = dup_axis(breaks = gdhi_regional_inc$monthly_value[which(gdhi_regional_inc$year == 2018 & 
                                                                                        gdhi_regional_inc$transaction == "Primary resources total")],
                                         labels = gdhi_regional_inc$sec_axis_labels[which(gdhi_regional_inc$year == 2018 & 
                                                                                          gdhi_regional_inc$transaction == "Primary resources total")],
                                         guide  = guide_axis(check.overlap = TRUE))) +
  scale_color_identity() +
  labs(title = "Average Regional Monthly Gross Income (1997 to 2018)",
       fill = NULL,
       x = NULL,
       y = NULL) +
  theme(plot.title = element_text(size = 24 , face = "bold" , hjust = 0, vjust = 7),
        plot.subtitle = element_text(size = 14 , hjust = 0 , vjust = 10),
        legend.position = c(0.18,1.02),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.key.width = unit(2.5, "cm"),
        panel.grid.minor = element_blank(), 
        panel.grid.major.x = element_blank(), 
        axis.title.y.right = element_blank(), 
        axis.text.y.right = element_text(size = 8),
        axis.text.x = element_text(size = 14, vjust = -2),
        plot.margin = margin(2,0.5,1,1, unit = "cm"))

# plot
plot_region_inc_timeseries

# write to directory as png
dir.create(here::here("projects/ons_region_gdhi_1997_2018/outputs"))

ggsave(plot = plot_region_inc_timeseries, here::here("projects/ons_region_gdhi_1997_2018/outputs/gdhi_inc_timeseries.png"), dpi = 360, height = 10, width = 16)

```

#### Map
```{r}
library(sf)

# Load geojson from url
uk_regions_boundaries <-
st_read('https://opendata.arcgis.com/datasets/01fd6b2d7600446d8af768005992f76a_3.geojson') %>%
  rename(region_name = nuts118nm) %>% 
  mutate(region_name = str_remove(region_name , " \\(England\\)"), 
         region_name = str_to_title(region_name))  %>%
  mutate(region_name = if_else(region_name == "Yorkshire And The Humber", "Yorkshire", region_name)) %>%
  mutate(region_name = if_else(region_name == "East Of England", "East of England", region_name))

# Merge for map
geo_gdhi_perhead_1997_2018 <- 
uk_regions_boundaries %>%
  inner_join(gdhi_regional_inc, by = "region_name")

# UK outline for map
library(rnaturalearth)
ireland <- ne_countries(scale = "medium", returnclass = "sf", country = "Ireland")

# Plot
plot_map_income_2018 <-
geo_gdhi_perhead_1997_2018 %>%
  filter(transaction == "Primary resources total") %>%
  filter(year == 2018) %>%
  ggplot() +
  geom_sf(data = ireland, fill = "white" , color = "grey60", alpha = 0.5) +
  geom_sf(aes(geometry = geometry, fill = value), color = "grey60" , alpha = 0.5, show.legend = FALSE) +
  #scale_fill_viridis_c(labels = scales::comma_format(accuracy = 1 , prefix = "£") , direction = -1 , begin = 0 , end = 0.8 , alpha = 0.6) +
  scale_fill_manual(values = palette) +
  coord_sf() +
  ggthemes::theme_map() +
  theme(plot.margin = margin(0,0,0,0, unit = "cm"))

# plot
plot_map_income_2018

# write to directory as png
dir.create(here::here("projects/ons_region_gdhi_1997_2018/outputs"))

ggsave(plot = plot_map_income_2018, here::here("projects/ons_region_gdhi_1997_2018/outputs/uk_primary_inc_map.png"), dpi = 360, height = 10, width = 8)

```

#### Panel
```{r}
library(cowplot)

x <- 0.8875

gdhi_ph_income_monthly <-
ggdraw() +
draw_plot(plot_map_income_2018, x = 0.2) +
draw_image(here::here("projects/ons_region_gdhi_1997_2018/outputs/gdhi_inc_timeseries.png")) +
#draw_text("Map Data: 2018 (current)" ,   size = 8 , hjust = 0 , x = 0.73 , y = 0.09, color = "grey60") +
draw_text("Data: ONS | Licence: Open Government Licence | Graphic: @NearAndDistant" ,   size = 12 , hjust = 0 , x = 0.0525 , y = 0.91) +
draw_text(gdhi_regional_inc$sec_axis_labels[which(gdhi_regional_inc$region_name == "West Midlands" & 
                                                  gdhi_regional_inc$year == 2018 &
                                                  gdhi_regional_inc$transaction == "Primary resources total")] , 
          size = 7.5 , hjust = 0 , x = x , y = 0.5225, color = "grey40") +
draw_text(gdhi_regional_inc$sec_axis_labels[which(gdhi_regional_inc$region_name == "Wales" & 
                                                  gdhi_regional_inc$year == 2018 &
                                                  gdhi_regional_inc$transaction == "Primary resources total")] , 
          size = 7.5 , hjust = 0 , x = x , y = 0.445, color = "grey40") +
theme(plot.background = element_rect(fill = "white" , color = "white"))

# Writing to File
ggsave(plot = gdhi_ph_income_monthly, here::here("projects/ons_region_gdhi_1997_2018/outputs/uk_gdhi_income_panel.png"), dpi = 360, height = 10, width = 16)

```

#########################
Primary Income Breakdown
#########################

#################################
Tables Indices (with/out London)

###### With London
```{r}
library(gt)
library(gtExtras)

tbl_inc_type_indi <-
gdhi_reg_inc_indi %>%
  filter(transaction %in% gdhi_income,
         transaction != "Primary resources total") %>%
  filter(year == 2018) %>%
  mutate(transaction = case_when(transaction == "Compensation of employees" ~ "Compensation",
                                 transaction == "Property income, received" ~ "Property income",
                                 TRUE ~ transaction),
         transaction = str_to_title(transaction)) %>%
  arrange(region_name) %>%
    pivot_wider(id_cols = c(region_name), names_from = transaction, values_from = value)

# add range - necessary to add into dataframe as {gt} cannot sum with missing values (i.e. London NA)
table_inc_indi_rng <- tbl_inc_type_indi %>% add_row(tbl_inc_type_indi %>% summarise(across(`Operating Surplus`:last_col(), ~max(.x, na.rm = TRUE) - min(.x, na.rm = TRUE))) %>% mutate(region_name = "Range"))

# actual table
table_inc_type_indi <-
table_inc_indi_rng %>%
  gt() %>%
      #### colour scale
      data_color(columns = c(`Operating Surplus`:`Property Income`),
                 # custom defined values - notice that order matters!
                 colors = scales::col_factor(palette = palette, domain = NULL), 
                 alpha = 0.4, autocolor_text = FALSE) %>% 
      tab_style(style = list(cell_fill(color = "white")), locations = cells_body(rows = 1)) %>%
      tab_style(style = list(cell_fill(color = "white")), locations = cells_body(rows = 14)) %>%
      #### cell boarder format
      cols_label(region_name = "") %>%
      cols_align(align = "left", columns = region_name) %>%
      cols_align(align = "center", columns = c(`Operating Surplus`:`Property Income`)) %>%
      tab_options(table.font.size = 16) %>%
      tab_style(style = list(cell_borders(sides = "bottom", color = "black", weight = px(2))),
                locations = list(cells_column_labels(columns = everything()))) %>%
      tab_style(style = list(cell_borders(sides = "top", color = "black", weight = px(2))),
                locations = list(cells_body(columns = everything(), rows = 14))) %>%
      tab_style(style = list(cell_text(align = "right")), locations = cells_body(rows = region_name == "Range")) %>%
      tab_source_note("Data: ONS (calculations own) | Table: @NearAndDistant") %>%
      tab_header(title = md("**2018 Gross Income Types as Indices;  UK Average = 100**")) %>%
      gt_theme_nytimes() 

table_inc_type_indi

```

```{r}

# Writing to File
gtsave_extra(data = table_inc_type_indi , 
             filename = here::here("projects/ons_region_gdhi_1997_2018/outputs/gdhi_inc_table_indi.png"), 
             zoom = 1.5 , vwidth = 700, vheight = 500)

```

###### Without London
```{r}
library(gt)
library(gtExtras)

tbl_inc_indi_wl <-
gdhi_regional_inc %>% 
  filter(transaction %in% gdhi_income,
         transaction != "Primary resources total") %>%
  filter(year == 2018) %>%
  #filter(region_name != "London") %>% # removal of London for this analysis
  mutate(transaction = case_when(transaction == "Compensation of employees" ~ "Compensation",
                                 transaction == "Property income, received" ~ "Property income",
                                 TRUE ~ transaction),
         transaction = str_to_title(transaction)) %>%
  arrange(region_name) %>%
  select(c(-region_ror, -sec_axis_labels, -monthly_value)) %>%
  # re-calculate indices without London
  pivot_wider(id_cols = c(region_name:year), names_from = region_name, values_from = value) %>%
  # Make London NA but keep in table
  mutate(`London` = NA) %>%
  rowwise() %>%
  mutate(`United Kingdom` = round(mean(`Scotland`:`Northern Ireland`, na.rm = TRUE))) %>%
  mutate(across(`Scotland`:`Northern Ireland`, ~ round(.x /`United Kingdom`,3)*100)) %>%
  mutate(`United Kingdom` = 100) %>%
  pivot_longer(cols = c(`United Kingdom`:`Northern Ireland`), names_to = "region_name") %>%
  # table generation
  pivot_wider(id_cols = region_name, names_from = transaction, values_from = value)

# add range - necessary to add into dataframe as {gt} cannot sum with missing values (i.e. London NA)
table_inc_indi_wl_rng <- tbl_inc_indi_wl %>% add_row(tbl_inc_indi_wl %>% summarise(across(`Operating Surplus`:last_col(), ~max(.x, na.rm = TRUE) - min(.x, na.rm = TRUE)))) %>% mutate(region_name = ifelse(is.na(region_name),"Range", region_name))

# actual table
table_inc_indi_wl <-
  table_inc_indi_wl_rng %>%
      gt() %>%
      #### colour scale
      data_color(columns = c(`Operating Surplus`:`Property Income`),
                 # custom defined values - notice that order matters!
                 colors = scales::col_factor(palette = palette, domain = NULL), 
                 alpha = 0.4, autocolor_text = FALSE) %>% 
      tab_style(style = list(cell_fill(color = "white")), locations = cells_body(rows = 1)) %>%
      # format missing London values
      fmt_missing(columns = everything(), missing_text = "-") %>%
      tab_style(style = list(cell_fill(color = "white")), locations = cells_body(rows = 11)) %>%
      tab_style(style = list(cell_fill(color = "white")), locations = cells_body(rows = 14)) %>%
      #### cell boarder format
      tab_options(table.font.size = 16) %>%
      cols_label(region_name = "") %>%
      cols_align(align = "left", columns = region_name) %>%
      cols_align(align = "center", columns = c(`Operating Surplus`:`Property Income`)) %>%
      tab_style(style = list(cell_borders(sides = "bottom", color = "black", weight = px(2))),
                locations = list(cells_column_labels(columns = everything()))) %>%
      tab_style(style = list(cell_borders(sides = "top", color = "black", weight = px(2))),
                locations = list(cells_body(columns = everything(), rows = 14))) %>%
      tab_style(style = list(cell_text(align = "right")), locations = cells_body(rows = region_name == "Range")) %>%
      tab_source_note(md("**Note:** Indices calculates without London for comparison")) %>%
      #tab_header(title = md("**2018 Gross Income Types as Indices of UK Average**")) %>%
      gt_theme_nytimes() 

table_inc_indi_wl

```

```{r}

# Writing to File
gtsave_extra(data = table_inc_indi_wl, 
             filename = here::here("projects/ons_region_gdhi_1997_2018/outputs/gdhi_inc_table_indi_wo_london.png"), 
             zoom = 1.5 , vwidth = 700, vheight = 500)

```

```{r}
library(cowplot)

panel_income_w_wo_lnd <-
ggdraw() +
  draw_image(here::here("projects/ons_region_gdhi_1997_2018/outputs/gdhi_inc_table_indi.png"), 
             height = 0.55 ,  width = 0.5 , x = 0.025, y = 0.2) +
  draw_image(here::here("projects/ons_region_gdhi_1997_2018/outputs/gdhi_inc_table_indi_wo_london.png"),
             height = 0.525 , width = 0.5 , x = 0.47, y = 0.1975) +
  theme(plot.margin = margin(-4,-2,-4,-2, unit = "cm"))

```

```{r}

# Writing to File
ggsave(plot = panel_income_w_wo_lnd, here::here("projects/ons_region_gdhi_1997_2018/outputs/tbl_inc_w_wo_lnd.png"), dpi = 360, height = 6, width = 12)

```


#######################
Income Type Range Plot
#######################

```{r}

gdhi_inc_ind_range <-
gdhi_reg_inc_indi %>% 
  select(-fill) %>%
  filter(transaction != "Primary resources total") %>%
  group_by(year , transaction) %>%
  summarise(range = (max(value) - min(value))/100) %>%
  ungroup() %>%
  group_by(transaction) %>%
  mutate(range_ror = last(range) - first(range)) %>%
  mutate(sec_axis_labels = paste0(transaction, " (+", scales::percent(range_ror, accuracy = 2),")"))

# sec_axis
sec_axis_range <-
gdhi_inc_ind_range %>%
  filter(year == 2018)

# plot
plot_inc_type_range <-
gdhi_inc_ind_range %>%
  ggplot(aes(year , range, color = transaction)) +
  geom_line(size = 1.5, show.legend = FALSE) +
  scale_x_continuous(breaks = c(1997,seq(2000,2015,5),2018), limits = c(1997,2018), expand = c(0,0)) +
  scale_y_continuous(labels = scales::percent_format(),
                     sec.axis = dup_axis(breaks = sec_axis_range$range,
                                         labels = sec_axis_range$sec_axis_labels)) +
  #ghibli::scale_fill_ghibli_d("MarnieLight1") +
  #nord::scale_color_nord(palette = "baie_mouton") +
  rcartocolor::scale_color_carto_d(palette = "Earth") +
  #scale_color_viridis_d(option = "magma") +
  #ochRe::scale_fill_ochre(palette = "dead_reef") +
  expand_limits(y = 0) +
  labs(x = NULL , y = NULL , color = NULL, title = "The Range Between Regions in the UK Across Different Income Types",
       subtitle = "Data: ONS | Licence: Open Government Licence | Graphic: @NearAndDistant") +
    theme(plot.title = element_text(size = 24 , face = "bold" , hjust = 0, vjust = 7),
        plot.subtitle = element_text(size = 14 , hjust = 0 , vjust = 10),
        legend.position = c(0.18,1.02),
        legend.direction = "horizontal",
        legend.title = element_blank(),
        legend.key.width = unit(2.5, "cm"),
        panel.grid.minor = element_blank(), 
        panel.grid.major.x = element_blank(), 
        axis.title.y.right = element_blank(), 
        axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size = 14, vjust = -2),
        plot.margin = margin(2,1,1,1, unit = "cm"))
  

```

```{r}

# Writing to File
ggsave(plot = plot_inc_type_range, here::here("projects/ons_region_gdhi_1997_2018/outputs/inc_type_range.png"), dpi = 360, height = 10, width = 16)

```

###########
UK Divided
###########

```{r}

# Plot
map_uk_div <-
cowplot::ggdraw(
gdhi_comp_indi_raw %>% 
  janitor::clean_names(numerals = "left") %>%
  filter(nuts_level %in% c("NUTS1" , "UK")) %>% # NUTS1 level is regions
  pivot_longer(cols = starts_with("x") , names_to = "year" , values_to = "value") %>%
  mutate(year = str_remove(year , "x"),
         year = as.numeric(year)) %>%
  mutate(region_name = if_else(region_name == "Yorkshire and The Humber", "Yorkshire", region_name)) %>%
  select(-nuts_level, -nuts_code, -transaction_code) %>%
  filter(transaction == "Gross Disposable Income") %>%
  filter(year == 2018) %>%
  left_join(uk_regions_boundaries, by = "region_name") %>%
  mutate(labels = paste0(region_name, ", ", value)) %>%
  mutate(value = value/100) %>%
  ggplot() +
  geom_sf(data = ireland, fill = "white" , color = "grey60", alpha = 0.5) +
  geom_sf(aes(geometry = geometry, fill = value), color = "grey60" , show.legend = TRUE) +
  ggrepel::geom_label_repel(aes(x = long, y = lat, label = labels), size = 3) +
  scale_fill_viridis_c(option = "viridis" , direction = 1 , alpha = 0.8,
                       breaks = seq(0.8,1.4,0.1), limits = c(0.8,1.4), labels=c("80%", "90%", "100% (UK Average)", "110%", "120%", "130%", "140%")) +
  labs(fill = NULL) +
  #ghibli::scale_fill_ghibli_c("PonyoMedium", direction = -1) +
  #nord::scale_fill_nord(palette = "baie_mouton", discrete = FALSE, reverse = TRUE) +
  #rcartocolor::scale_fill_carto_c(palette = "TealRose", direction = -1) +
  #scale_color_viridis_d(option = "magma") +
  #ochRe::scale_fill_ochre(palette = "galah", discrete = FALSE, ) +
  coord_sf() +
  ggthemes::theme_map() +
  theme(plot.margin = margin(0,0,0,0, unit = "cm"),
        legend.position = c(0.75,0.55),
        legend.direction = "vertical",
        legend.title = element_blank(),
        legend.key.width = unit(1, "cm"))) +
draw_text("2018 Regional Differences in\nDisposable Income, as Indices\nUK Average = 100",
           x = 0.14, y = 0.13, hjust = 0 , size = 12, fontface = "bold") +
draw_text("Data: ONS\nGraphic: @NearAndDistant",
           x = 0.14, y = 0.06, hjust = 0 , size = 9)

```

```{r}

# Writing to File
ggsave(plot = map_uk_div, here::here("outputs/map_uk_div.png"), dpi = 360, height = 10, width = 8, bg = "white")

```

##############
Housing Costs
##############

```{r}

tbl_housing_costs <- 
gdhi_reg_inc_indi_cln %>%
  filter(transaction == "Property income, paid") %>%
  arrange(region_name) %>%
  group_by(region_name) %>%
  mutate(trend = list(value)) %>%
  ungroup() %>%
    pivot_wider(id_cols = c(region_name, trend), names_from = year, values_from = value) %>%
      gt(rowname_col = "region_name") %>%
      gt_sparkline(trend, width = 25) %>%
      grand_summary_rows(columns = c(`1997`:`2018`), fns = list(Range = ~max(.) - min(.)), use_seps = TRUE, missing_text = "") %>%
      #### color scale
      data_color(columns = c(`1997`:`2018`),
                 # custom defined values - notice that order matters!
                 colors = scales::col_factor(palette = rev(palette), domain = NULL), 
                 alpha = 0.4, autocolor_text = FALSE) %>% 
      tab_style(style = list(cell_fill(color = "white")), locations = cells_body(rows = 1)) %>%
      #### cell boarder format
      tab_style(style = list(cell_borders(sides = "right", color = "black", weight = px(3))),
                locations = list(cells_body(columns = trend))) %>%
      tab_style(style = list(cell_borders(sides = "top", color = "grey60", weight = px(1))),
                locations = list(cells_column_labels(columns = everything()))) %>%
      tab_style(style = list(cell_borders(sides = "bottom", color = "black", weight = px(3))),
                locations = list(cells_column_labels(columns = everything()))) %>%
      tab_source_note("Data: ONS | Table: @NearAndDistant") %>%
      tab_header(title = md("**Property Payments (Rent & Mortgage) from 1997 to 2018 as Indices of UK Average; UK Average = 100**")) %>%
      gt_theme_nytimes() 

```

```{r}

# Writing to File
gtsave_extra(data = tbl_housing_costs, filename = here::here("projects/ons_region_gdhi_1997_2018/outputs/tbl_housing costs.png"), 
             zoom = 1.5 , vwidth = 1400, vheight = 500)

```
