---
title: "NHS-R Sankey"
author: "NearAndDistant"
date: "22/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Grab ONS Gross Data from GitHub (Script in Repository)

Gross Disposable household Income is the amount of money that that all of the individuals in the household sector have available for spending or saving after income distribution measures (for example, taxes, social contributions and benefits) have taken effect. GDHI does not provide measures relating to actual households or family units. The figures cover regions, sub-regions and local areas of the UK. It consists of:

##### Primary resources
0: Operating surplus
1: Mixed income
2: Compensation of employees
3: Property income, received

= Primary Resource total

##### Primary uses
4: Property income, paid

##### Secondary resources
5: Social benefits received
6: Other current transfers, received

= Balance of Primary Income

##### Secondary uses
7: Current taxes on income, wealth etc
8: Social contributions, paid
9: Other current transfers, paid

##### GDHI
= Gross Disposable Income

```{r}
library(tidyverse)

# you can see the primary script to grab this at https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gross_income_1997_to_2018/
ons_gdhi_components_97_18_raw <-
read_csv("https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gdhi_97_18/ons_gdhi_components_97_18.csv")

```

##### Grabs NUTS1
```{r}

ons_gdhi_NUTS1 <-
ons_gdhi_components_97_18_raw %>%
  filter(`NUTS level` == "NUTS1") %>% # NUTS1 level is regions
  janitor::clean_names(numerals = "left") %>%
  pivot_longer(cols = starts_with("x") , names_to = "year" , values_to = "value") %>%
  mutate(year = str_remove(year , "x"),
         year = as.numeric(year),
         region_name = str_to_title(region_name)) %>%
  # clean long names
  mutate(region_name = if_else(region_name == "Yorkshire And The Humber", "Yorkshire", region_name)) %>%
  mutate(transaction = if_else(transaction == "Imputed social contributions/Social benefits received" , "Social benefits received", transaction)) %>%
  mutate(transaction = if_else(transaction == "Social contributions/Social benefits paid" , "Social contributions, paid", transaction)) %>%
    mutate(transaction = if_else(transaction == "Current taxes on income, wealth etc" , "Current taxes", transaction)) %>%
  # filter out unecessary transactions i.e. totals or balances
  mutate(transaction_code = if_else(transaction == "Primary resources total" , "PRT", transaction_code)) %>% # so we keep this as a flow node
  mutate(value = if_else(transaction == "Gross Disposable Income" , NA_real_, value)) %>% # so we keep this as a flow node
  drop_na(transaction_code) %>% # NAs are just total rows in the ONS Excel file
  filter(year == 2018) %>%
  # testing filter (two regions)
  #filter(region_name == "North East" | region_name == "London") %>% 
  select(-year, -nuts_level, -nuts_code, -transaction_code) %>%
  # balancing out financials: as the flow diagram needs to know what is left over after "secondary payments" are mad (taxes etc)
  # we have to calculate their subtraction from Balance of Payments so that Gross Dispo
  pivot_wider(id_cols = c(region_name:transaction), names_from = "transaction", values_from = "value") %>%
  mutate(`Balance of primary incomes` = `Balance of primary incomes` - `Current taxes` - `Social contributions, paid` - `Other current transfers, paid`) %>%
  pivot_longer(cols = c(`Operating surplus`:`Gross Disposable Income`), names_to = "transaction" , values_to = "value") %>%
  mutate(transaction = if_else(transaction == "Balance of primary incomes" , "Secondary resources total", transaction)) # aligning to the ONS figures on Secondary Income

```

##### Adding Group
```{r}
prep_NUTS1 <-
ons_gdhi_NUTS1 %>%
mutate(
source = 
case_when(
transaction == "Operating surplus"                   ~ "Operating surplus",
transaction == "Mixed income"                        ~ "Mixed income",
transaction == "Compensation of employees"           ~ "Compensation of employees",
transaction == "Property income, received"           ~ "Property income, received",
transaction == "Primary resources total"             ~ "Primary resources total",
transaction == "Property income, paid"               ~ "Primary resources total",
transaction == "Secondary resources total"           ~ "Secondary resources total",
transaction == "Social benefits received"            ~ "Social benefits received",
transaction == "Other current transfers, received"   ~ "Other current transfers, received",
transaction == "Current taxes"                       ~ "Secondary resources total",
transaction == "Social contributions, paid"          ~ "Secondary resources total",
transaction == "Other current transfers, paid"       ~ "Secondary resources total",
transaction == "Gross Disposable Income"             ~ "Secondary resources total"),
target = 
case_when(
transaction  == "Operating surplus"                   ~ "Primary resources total",
transaction  == "Mixed income"                        ~ "Primary resources total",
transaction  == "Compensation of employees"           ~ "Primary resources total",
transaction  == "Property income, received"           ~ "Primary resources total",
transaction  == "Primary resources total"             ~ "Secondary resources total",
transaction  == "Property income, paid"               ~ "Property income, paid",
transaction  == "Secondary resources total"           ~ "Gross Disposable Income",
transaction  == "Social benefits received"            ~ "Secondary resources total",
transaction  == "Other current transfers, received"   ~ "Secondary resources total",
transaction  == "Current taxes"                       ~ "Current taxes",
transaction  == "Social contributions, paid"          ~ "Social contributions, paid",
transaction  == "Other current transfers, paid"       ~ "Other current transfers, paid",
transaction  == "Gross Disposable Income"             ~ "Gross Disposable Income")) %>%
select(region_name , source , target , value) %>%
group_by(region_name)

# create line ids
indexed_NUTS1 <-
prep_NUTS1 %>%
  group_by(region_name) %>%
  mutate(id = str_pad(7:19, width = 2, pad = "0")) %>%
  select(id , everything()) %>%
  ungroup()

```

```{r}

regions <- 
  tibble(
    id          = indexed_NUTS1$id,
    region_name = indexed_NUTS1$region_name,
    source      = indexed_NUTS1$region_name,
    target      = indexed_NUTS1$source,
    value       = indexed_NUTS1$value) %>%
    group_by(source) %>%
    filter(id %in% c("07", "08" , "09" , "10" , "14" , "15")) %>%
    # create new ids for 01 - 06
    group_by(region_name) %>%   
    mutate(id = str_pad(1:6, width = 2, pad = "0")) %>%
    select(id , everything())

# join
full_nuts1 <-
  rbind(indexed_NUTS1 , regions)

# create group ids additional to line ids to force sankey to order itself North to South
full_nuts1 <-
full_nuts1 %>%
mutate(
id = 
case_when(
region_name  == "Scotland"         ~ paste("01" , id, sep = "."),
region_name  == "North East"       ~ paste("02" , id, sep = "."),
region_name  == "North West"       ~ paste("03" , id, sep = "."),
region_name  == "Yorkshire"        ~ paste("04" , id, sep = "."),
region_name  == "East Midlands"    ~ paste("05" , id, sep = "."),
region_name  == "West Midlands"    ~ paste("06" , id, sep = "."),
region_name  == "Wales"            ~ paste("07" , id, sep = "."),
region_name  == "South West"       ~ paste("08" , id, sep = "."),
region_name  == "South East"       ~ paste("09" , id, sep = "."),
region_name  == "London"           ~ paste("10" , id, sep = "."),
region_name  == "East Of England"  ~ paste("11", id, sep = "."),
region_name  == "Northern Ireland" ~ paste("12", id, sep = "."))) %>%
arrange(id)

```

#### Create networkD3 links

Great reference on this by NHS-R (Lisa Cummmins), here:
https://www.youtube.com/watch?v=3OvsW8OI1wo

```{r}

# Nodes all
nodes <- 
  indexed_NUTS1 %>%
  select(source , target) %>%
  pivot_longer(cols = c(source,target), values_to = "table_link") %>%
  select(-name) %>%
  distinct(table_link, .keep_all = TRUE)

# regions *in order* bind
regions_ordered <-
  tibble(
  table_link = 
  unique(full_nuts1$region_name))

# join for all sources giving all nodes and region nodes in order
nodes_all <- rbind(regions_ordered, nodes)

```

#### Create node links - All Regions
```{r}

networkD3_table <-
  full_nuts1 %>%
  mutate(source_id  = match(source , nodes_all$table_link)-1) %>%
  mutate(target_id = match(target ,  nodes_all$table_link)-1)

```

#### Palette

```{r}
# color scales
library(ggsci)

scales::show_col(pal_rickandmorty()(12))
pal_rickandmorty()(12)

scales::show_col(pal_futurama()(12))

## viridis

scales::show_col(viridis::cividis(n = 12))
viridis::cividis(n = 12) # return string of hex colors to create palette

```

NUTS1 Palette
"#FFFFFF" , "#e2d8f4", "#ffefd9", "#dff3f5", "#efead7", "#ffffbf", "#bfc9ad", "#d1ee70", "#ffc9c9", "#66e6d9", "#dff2bd", "#fcc0a6", "#68B374"

Viridis
"#FFFFFF", "#440154FF", "#482173FF", "#433E85FF", "#38598CFF", "#2D708EFF", "#25858EFF", "#1E9B8AFF", "#2BB07FFF", "#51C56AFF", "#85D54AFF", "#C2DF23FF", "#FDE725FF"

Plasma
"#FFFFFF", "#0D0887FF", "#3E049CFF", "#6300A7FF", "#8707A6FF", "#A62098FF", "#C03A83FF", "#D5546EFF", "#E76F5AFF", "#F58C46FF", "#FDAD32FF", "#FCD225FF", "#F0F921FF"

Mako
"#FFFFFF" , "#0B0405FF", "#231526FF", "#35264CFF", "#403A75FF", "#3D5296FF", "#366DA0FF", "#3487A6FF", "#35A1ABFF", "#43BBADFF", "#6CD3ADFF", "#ADE3C0FF", "#DEF5E5FF"

Turbo
"#FFFFFF", "#30123BFF",  "#4454C4FF", "#4490FEFF", "#1FC8DEFF", "#29EFA2FF", "#7DFF56FF", "#C1F334FF", "#F1CA3AFF", "#FE922AFF", "#EA4F0DFF", "#BE2102FF", "#7A0403FF"

Cividis
"#FFFFFF", "#00204DFF", "#00306FFF", "#2A406CFF", "#48526BFF", "#5E626EFF", "#727374FF", "#878479FF", "#9E9677FF", "#B6A971FF", "#D0BE67FF", "#EAD357FF", "#FFEA46FF"

```{r}
# map colors to dataframe

# All Regions
networkD3_table <-
networkD3_table %>%
  mutate(
    color_group = case_when(
    region_name == "Scotland"         ~ "Purple"     ,
    region_name == "North East"       ~ "Orange"     ,
    region_name == "North West"       ~ "Blue"       ,
    region_name == "Yorkshire"        ~ "Brown"      ,
    region_name == "East Midlands"    ~ "Yellow"     ,
    region_name == "West Midlands"    ~ "Olive"      ,
    region_name == "Wales"            ~ "Lime"       ,
    region_name == "South West"       ~ "Pink"       ,
    region_name == "South East"       ~ "Grey"       ,
    region_name == "London"           ~ "Green"      ,
    region_name == "East Of England"  ~ "Tangerine"  ,
    region_name == "Northern Ireland" ~ "Clover"))

# For some reason JavaScript does not take spaces in it's characters indicating colors so use one word colors otherwise it will skip the color and place the next. Use one word, CamelCase or snake_case.

link_color <- 
htmlwidgets::JS('d3.scaleOrdinal()
    .domain(["White", "Purple","Orange","Blue","Brown","Yellow","Olive","Lime","Pink","Grey","Green","Tangerine","Clover" ])
    .range(["#FFFFFF" , "#9E7B99", "#B7E4F9B3", "#24325F99", "#82491E99", "#FB646799", "#FF634899", "#ADE2D099", "#E8924299", "#FAE48B99", "#3F404199", "#917C5D99", "#526E2D99"])')

# opacity of hex colors: https://gist.github.com/lopspower/03fb1cc0ac9f32ef38f4
```

#### {networkD3} All Regions
```{r}
library(networkD3)

nd3_gdhi <- 
list(
  nodes = data.frame(name   =        nodes_all$table_link,
                     color  =        "will pick first color i.e. white"),
  links = data.frame(source =        networkD3_table$source_id,
                     target =        networkD3_table$target_id,
                     value  =        networkD3_table$value,
                     color  =        networkD3_table$color_group))


gdhi_nuts1_sankey <-
sankeyNetwork(Links = nd3_gdhi$links, Nodes = nd3_gdhi$nodes , 
              Source = "source", Target = "target", Value = "value", 
              LinkGroup = "color", colourScale = link_color,
              NodeID = "name", NodeGroup = "color", nodeWidth = 30,
              units = "(£m)", fontSize = 16, fontFamily = "Arial" ,
              width = 1700 , height = 900 ,  sinksRight = FALSE, iterations = 0,
              margin = list(right = -40, left = 250))

# render with js
gdhi_nuts1_sankey <- 
htmlwidgets::onRender(gdhi_nuts1_sankey,
  'function(el, x) {
    d3.selectAll(".node text").attr("fill", "black");
    d3.selectAll(".node text").attr("text-anchor", "end").attr("x", -10);
    d3.selectAll(".node text").attr("fill", "black");
    d3.select("body").style("background-color", "#FFFFFF"); 
  }'
) 

gdhi_nuts1_sankey

```

#### Saving
```{r}

# create directory
dir.create("projects/ons_region_gdhi_1997_2018/outputs")

dir.create("projects/ons_region_gdhi_1997_2018/outputs/html")
dir.create("projects/ons_region_gdhi_1997_2018/outputs/images")

# save it as html
saveNetwork(gdhi_nuts1_sankey, here::here("projects/ons_region_gdhi_1997_2018/outputs/images/gdhi_sankey_ggsci.html"))

# convert it as png
webshot2::webshot(here::here("projects/ons_region_gdhi_1997_2018/outputs/images/gdhi_sankey_ggsci.html"),
                  here::here("projects/ons_region_gdhi_1997_2018/outputs/images/gdhi_sankey_ggsci.png"),
                  vwidth = 1500, vheight = 800)

```

#### Map Legend

###### Create UK Outline
```{r}
#### grab national outline
library(rnaturalearth)

uk      <- ne_countries(scale = "medium", returnclass = "sf", country = "United Kingdom")
ireland <- ne_countries(scale = "medium", returnclass = "sf", country = "Ireland")

# grab boundaries from ONS
library(sf)

# load geojson from url
uk_regions_boundaries <-
st_read('https://opendata.arcgis.com/datasets/01fd6b2d7600446d8af768005992f76a_3.geojson') %>%
  rename(region_name = nuts118nm) %>% 
  mutate(region_name = str_remove(region_name , " \\(England\\)"), 
         region_name = str_to_title(region_name))  %>%
  mutate(region_name = if_else(region_name == "Yorkshire And The Humber", "Yorkshire", region_name))

# create tibble for colors
nuts1_palette <- tibble(region_name = c("Scotland", "North East", "North West", "Yorkshire", "East Midlands", "West Midlands",
                                   "Wales", "South West", "South East", "London", "East Of England", "Northern Ireland"),
                        fill  = c("#9E7BB5", "#B7E4F9FF", "#24325FFF", "#82491EFF", "#FB6467FF", "#FF6348FF", 
                                  "#ADE2D0FF", "#E89242FF", "#FAE48BFF", "#3F4041FF", "#917C5DFF", "#526E2DFF"))

# join
uk_regions_boundaries <-
  uk_regions_boundaries %>%
  left_join(nuts1_palette , by = "region_name")

```

##### Plot Map
```{r}
library(ggrepel)

plot_region_legend <-
uk_regions_boundaries %>%
  ggplot() +
  geom_sf(data = ireland, fill = "white" , color = "grey80") +
  geom_sf(aes(fill = fill), alpha = 0.75 , color = "white" , show.legend = FALSE) +
  geom_label_repel(aes(label = region_name, x = long, y = lat)) +
  scale_fill_identity(guide = "legend") +
  coord_sf() +
  ggthemes::theme_map()

ggsave(here::here("projects/ons_region_gdhi_1997_2018/outputs/images/ggsci_palette.png"), dpi = 360, height = 9.5, width = 15)

```

#### Annotation
```{r}
library(cowplot)

ggdraw() +
  draw_image(here::here("projects/ons_region_gdhi_1997_2018/outputs/images/gdhi_sankey_ggsci.png"), height = 0.95, width = 1.5, x = -0.27, y = 0.07) +
  draw_image(here::here("projects/ons_region_gdhi_1997_2018/outputs/images/ggsci_palette.png"), height = 0.8, x = 0.34, y = -0.05) +
  draw_text("UK Regional Gross Disposable Household Income (GDHI), 2018" , hjust = 0 , x = 0.01 , y = 0.04, family = "Arial", color = "black", size = 20) +
  draw_text("Data: ONS | Graphic: @NearAndDistant" , hjust = 0 , x = 0.01 , y = 0.07, family = "Arial", color = "black", size = 12)


ggsave(here::here("projects/ons_region_gdhi_1997_2018/outputs/images/ggsci_palette.jpeg"), dpi = 300, height = 10, width = 16)

```
