---
title: "NHS-R Sankey"
author: "NearAndDistant"
date: "22/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Grab ONS Gross Data from GitHub (Script in Repository)

Gross Disposable household Income is the amount of money that that all of the individuals in the household sector have available for spending or saving after income distribution measures (for example, taxes, social contributions and benefits) have taken effect. GDHI does not provide measures relating to actual households or family units. The figures cover regions, sub-regions and local areas of the UK. It consists of:

##### Primary resources
0: Operating surplus
1: Mixed income
2: Compensation of employees
3: Property income, received

= Primary Resource total

##### Primary uses
4: Property income, paid

##### Secondary resources
5: Social benefits received
6: Other current transfers, received

= Balance of Primary Income

##### Secondary uses
7: Current taxes on income, wealth etc
8: Social contributions, paid
9: Other current transfers, paid

##### GDHI
= Gross Disposable Income

```{r}
library(tidyverse)

# you can see the primary script to grab this at https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gross_income_1997_to_2018/
ons_gdhi_components_97_18_raw <-
read_csv("https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gdhi_97_18/ons_gdhi_components_97_18.csv")

```

##### Grabs NUTS1
```{r}

ons_gdhi_NUTS1 <-
ons_gdhi_components_97_18_raw %>%
  filter(`NUTS level` == "NUTS1") %>% # NUTS1 level is regions
  janitor::clean_names(numerals = "left") %>%
  pivot_longer(cols = starts_with("x") , names_to = "year" , values_to = "value") %>%
  mutate(year = str_remove(year , "x"),
         year = as.numeric(year),
         region_name = str_to_title(region_name)) %>%
  # clean long names
  mutate(region_name = if_else(region_name == "Yorkshire And The Humber", "Yorkshire", region_name)) %>%
  mutate(transaction = if_else(transaction == "Imputed social contributions/Social benefits received" , "Social benefits received", transaction)) %>%
  mutate(transaction = if_else(transaction == "Social contributions/Social benefits paid" , "Social contributions, paid", transaction)) %>%
  # filter out unecessary transactions i.e. totals or balances
  mutate(transaction_code = if_else(transaction == "Primary resources total" , "PRT", transaction_code)) %>% # so we keep this as a flow node
  mutate(value = if_else(transaction == "Gross Disposable Income" , NA_real_, value)) %>% # so we keep this as a flow node
  drop_na(transaction_code) %>% # NAs are just total rows in the ONS Excel file
  filter(year == 2018) %>%
  # testing filter (two regions)
  #filter(region_name == "North East" | region_name == "London") %>% 
  select(-year, -nuts_level, -nuts_code, -transaction_code) %>%
  # balancing out financials: as the flow diagram needs to know what is left over after "secondary payments" are mad (taxes etc)
  # we have to calculate their subtraction from Balance of Payments so that Gross Dispo
  pivot_wider(id_cols = c(region_name:transaction), names_from = "transaction", values_from = "value") %>%
  mutate(`Balance of primary incomes` = `Balance of primary incomes` - `Current taxes on income, wealth etc` - `Social contributions, paid` - `Other current transfers, paid`) %>%
  pivot_longer(cols = c(`Operating surplus`:`Gross Disposable Income`), names_to = "transaction" , values_to = "value")

```

##### Adding Group
```{r}
prep_NUTS1 <-
ons_gdhi_NUTS1 %>%
mutate(
source = 
case_when(
transaction == "Operating surplus"                   ~ "Operating surplus",
transaction == "Mixed income"                        ~ "Mixed income",
transaction == "Compensation of employees"           ~ "Compensation of employees",
transaction == "Property income, received"           ~ "Property income, received",
transaction == "Primary resources total"             ~ "Primary resources total",
transaction == "Property income, paid"               ~ "Balance of primary incomes",
transaction == "Balance of primary incomes"          ~ "Balance of primary incomes",
transaction == "Social benefits received"            ~ "Social benefits received",
transaction == "Other current transfers, received"   ~ "Other current transfers, received",
transaction == "Current taxes on income, wealth etc" ~ "Balance of primary incomes",
transaction == "Social contributions, paid"          ~ "Balance of primary incomes",
transaction == "Other current transfers, paid"       ~ "Balance of primary incomes",
transaction == "Gross Disposable Income"             ~ "Balance of primary incomes"),
target = 
case_when(
transaction  == "Operating surplus"                   ~ "Primary resources total",
transaction  == "Mixed income"                        ~ "Primary resources total",
transaction  == "Compensation of employees"           ~ "Primary resources total",
transaction  == "Property income, received"           ~ "Primary resources total",
transaction  == "Primary resources total"             ~ "Balance of primary incomes",
transaction  == "Property income, paid"               ~ "Property income, paid",
transaction  == "Balance of primary incomes"          ~ "Gross Disposable Income",
transaction  == "Social benefits received"            ~ "Balance of primary incomes",
transaction  == "Other current transfers, received"   ~ "Balance of primary incomes",
transaction  == "Current taxes on income, wealth etc" ~ "Current taxes on income, wealth etc",
transaction  == "Social contributions, paid"          ~ "Social contributions, paid",
transaction  == "Other current transfers, paid"       ~ "Other current transfers, paid",
transaction  == "Gross Disposable Income"             ~ "Gross Disposable Income")) %>%
select(region_name , source , target , value) %>%
group_by(region_name)

# create line ids
indexed_NUTS1 <-
prep_NUTS1 %>%
  group_by(region_name) %>%
  mutate(id = str_pad(7:19, width = 2, pad = "0")) %>%
  select(id , everything()) %>%
  ungroup()

```

##### Adding Regions
```{r}

regions <- 
  tibble(
    id          = indexed_NUTS1$id,
    region_name = indexed_NUTS1$region_name,
    source      = indexed_NUTS1$region_name,
    target      = indexed_NUTS1$source,
    value       = indexed_NUTS1$value) %>%
    group_by(source) %>%
    filter(id %in% c("07", "08" , "09" , "10" , "14" , "15")) %>%
    # create new ids for 01 - 06
    group_by(region_name) %>%   
    mutate(id = str_pad(1:6, width = 2, pad = "0")) %>%
    select(id , everything())

# join
full_nuts1 <-
  rbind(indexed_NUTS1 , regions)

# create group ids additional to line ids to force sankey to order itself North to South
full_nuts1 <-
full_nuts1 %>%
mutate(
id = 
case_when(
region_name  == "Scotland"         ~ paste("01" , id, sep = "."),
region_name  == "North East"       ~ paste("02" , id, sep = "."),
region_name  == "North West"       ~ paste("03" , id, sep = "."),
region_name  == "Yorkshire"        ~ paste("04" , id, sep = "."),
region_name  == "East Midlands"    ~ paste("05" , id, sep = "."),
region_name  == "West Midlands"    ~ paste("06" , id, sep = "."),
region_name  == "Wales"            ~ paste("07" , id, sep = "."),
region_name  == "South West"       ~ paste("08" , id, sep = "."),
region_name  == "South East"       ~ paste("09" , id, sep = "."),
region_name  == "London"           ~ paste("10" , id, sep = "."),
region_name  == "East Of England"  ~ paste("11", id, sep = "."),
region_name  == "Northern Ireland" ~ paste("12", id, sep = "."))) %>%
arrange(id)

```

Great reference on this by NHS-R (Lisa Cummmins), here:
https://www.youtube.com/watch?v=3OvsW8OI1wo

#### Create networkD3 links
```{r}

# Nodes
nodes <- 
  indexed_NUTS1 %>%
  select(source , target) %>%
  pivot_longer(cols = c(source,target), values_to = "table_link") %>%
  select(-name) %>%
  distinct(table_link)

# regions *in order* bind
regions_ordered <-
  tibble(
  table_link = 
  unique(full_nuts1$region_name))

# join for all sources giving all nodes and region nodes in order
nodes <-
  rbind(regions_ordered, nodes)

```

#### Create node links
```{r}

networkD3_table <-
  full_nuts1 %>%
  mutate(source_id  = match(source , nodes$table_link)-1) %>%
  mutate(target_id = match(target , nodes$table_link)-1)

```

#### Node colors

"Scotland"          = "Purple"          = "#e2d8f4"       
"North East"        = "Orange"          = "#ffefd9"        
"North West"        = "Blue"            = "#dff3f5"        
"Yorkshire"         = "Brown"           = "#efead7"        
"East Midlands"     = "Yellow"          = "#ffffbf"            
"West Midlands"     = "Olive"           = "#bfc9ad"           
"Wales"             = "Lime"            = "#d1ee70"        
"South West"        = "Pink"            = "#ffc9c9"         
"South East"        = "Grey"            = "#bfc9c7"         
"London"            = "Green"           = "#dff2bd"        
"East Of England"   = "Tangerine"       = "#fcc0a6"              
"Northern Ireland"  = "Clover"          = "#68B374"   

##### Show colors
```{r}

nuts1_palette <- c("#e2d8f4", "#ffefd9", "#dff3f5", "#efead7", "#ffffbf", "#bfc9ad", "#d1ee70", "#ffc9c9", "#bfc9c7", "#dff2bd", "#fcc0a6", "#68B374")

scales::show_col(nuts1_palette)

```

For some reason JavaScript does not take spaces in it's characters indicating colors so use one word colors otherwise it will skip the color and place the next. Use one word, CamelCase or snake_case.

```{r}

# color scales

## viridis

viridis::viridis(n = 12)
scales::show_col(viridis::viridis(n = 12))

networkD3_table <-
networkD3_table %>%
  mutate(
    color_group = case_when(
    region_name == "Scotland"         ~ "Purple"     ,
    region_name == "North East"       ~ "Orange"     ,
    region_name == "North West"       ~ "Blue"       ,
    region_name == "Yorkshire"        ~ "Brown"      ,
    region_name == "East Midlands"    ~ "Yellow"     ,
    region_name == "West Midlands"    ~ "Olive"      ,
    region_name == "Wales"            ~ "Lime"       ,
    region_name == "South West"       ~ "Pink"       ,
    region_name == "South East"       ~ "Grey"       ,
    region_name == "London"           ~ "Green"      ,
    region_name == "East Of England"  ~ "Tangerine"  ,
    region_name == "Northern Ireland" ~ "Clover"))

link_color <- 
JS('d3.scaleOrdinal()
    .domain(["White", "Purple","Orange","Blue","Brown","Yellow","Olive","Lime","Pink","Grey","Green","Tangerine","Clover" ])
    .range(["#FFFFFF","#440154FF", "#482173FF", "#433E85FF", "#38598CFF", "#2D708EFF", "#25858EFF", "#1E9B8AFF", "#2BB07FFF", "#51C56AFF", "#85D54AFF", "#C2DF23FF", "#FDE725FF"])')



```

"#e2d8f4","#ffefd9","#dff3f5","#efead7","#ffffbf","#bfc9ad","#d1ee70","#ffc9c9","#bfc9c7","#dff2bd","#fcc0a6","#68B374"

#### {networkD3}
```{r}
library(networkD3)

nd3_gdhi <- 
list(
  nodes = data.frame(name   =        nodes$table_link,
                     color  =        "will pick first color i.e. white"),
  links = data.frame(source =        networkD3_table$source_id,
                     target =        networkD3_table$target_id,
                     value  =        networkD3_table$value,
                     color  =        networkD3_table$color_group))


gdhi_nuts1_sankey <-
sankeyNetwork(Links = nd3_gdhi$links, Nodes = nd3_gdhi$nodes , 
              Source = "source", Target = "target", Value = "value", 
              LinkGroup = "color", colourScale = link_color,
              NodeID = "name", NodeGroup = "color", nodeWidth = 30,
              units = "(£m)", fontSize = 14, fontFamily = "Arial" , 
              width = 1500 , height = 900 ,  sinksRight = FALSE, iterations = 0,
              linkDistance = JS('function(){d3.select("body").style("background-color", "#DAE3F9"); return 50;}')) # hack to change background color

# render with JavaScript to change characteristics
# render with js
gdhi_nuts1_sankey <- 
htmlwidgets::onRender(gdhi_nuts1_sankey,
  'function(el, x) {
    d3.selectAll(".node text")
        .style("fill", "#444444");
  }'
)

gdhi_nuts1_sankey

```

#### Saving
```{r}

# create directory
dir.create("projects/ons_region_gdhi_1997_2018/outputs")

# save it as html
saveNetwork(gdhi_nuts1_sankey, here::here("projects/ons_region_gdhi_1997_2018/outputs/gdhi_nuts1_sankey.html"))

library(webshot)
# you convert it as png
webshot(here::here("projects/ons_region_gdhi_1997_2018/outputs/gdhi_nuts1_sankey.html"),
        here::here("projects/ons_region_gdhi_1997_2018/outputs/gdhi_nuts1_sankey.png"), vwidth = 1200, vheight = 900)

```

