---
title: "ONS GDHI Components"
author: "NearAndDistant"
date: "19/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Grab ONS Gross Data from GitHub (Script in Repository)
```{r}
library(tidyverse)

# you can see the primary script to grab this at https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gross_income_1997_to_2018/
ons_gdhi_components_97_18 <-
read_csv("https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gdhi_components/ons_gdhi_components_97_18.csv")

```

##### Grabs NUTS1
```{r}

ons_gdhi_components_97_18_NUTS1 <-
ons_gdhi_components_97_18 %>%
  filter(`NUTS level` == "NUTS1") %>%
  janitor::clean_names(numerals = "left") %>%
  pivot_longer(cols = starts_with("x") , names_to = "year" , values_to = "value") %>%
  drop_na(transaction_code) %>% # NAs are just total rows in the ONS Excel file
  filter(transaction != "Balance of primary incomes") %>% # not applicable as simple total rows in the ONS Excel file
  mutate(year = str_remove(year , "x"),
         year = as.numeric(year),
         region_name = str_to_title(region_name)) %>%
  mutate(region_name = if_else(region_name == "Yorkshire And The Humber", "Yorkshire", region_name)) %>%
  select(year , everything())

```

#### Prepare Sankey

##### Primary resources
0: Operating surplus
1: Mixed income
2: Compensation of employees
3: Property income, received

##### Primary uses
4: Property income, paid

##### Secondary resources
5: Social benefits received
6: Other current transfers, received

##### Secondary uses
7: Current taxes on income, wealth etc
8: Social contributions/Social benefits paid
9: Other current transfers, paid

##### GDHI
10: Gross Disposable Income

```{r}

nd3_flow <- 
tibble(
ref   = 0:15,
source = c(
"Operating surplus",                      
"Mixed income",              
"Compensation of employees",                 
"Property income, received",              
"Property income, paid",  
"Property income, paid",
"Social benefits received",
"Social benefits received",
"Social benefits received",
"Other current transfers, received",
"Other current transfers, received",
"Other current transfers, received",
"Current taxes on income, wealth etc",      
"Social contributions, paid",
"Other current transfers, paid",
"Gross Disposable Income"),
target = c(
"Property income, paid",
"Property income, paid",
"Property income, paid",
"Property income, paid",
"Social benefits received",
"Other current transfers, received",
"Current taxes on income, wealth etc",      
"Social contributions, paid",
"Other current transfers, paid",
"Current taxes on income, wealth etc",      
"Social contributions, paid",
"Other current transfers, paid",
"Gross Disposable Income",
"Gross Disposable Income",
"Gross Disposable Income",
NA
))

nd3_flow <-
tibble(
  source = as.integer(c(0,1,2,3,4,4,5,5,5,6,6,6,7, 8, 9)),
  target = as.integer(c(4,4,4,4,5,6,7,8,9,7,8,9,10,10,10))
)

```

##### Gross Disposable Household Income (GDHI)

```{r}
library(ggsankey)

gdhi_sankey_prep <-
ons_gdhi_components_97_18_NUTS1 %>%
  mutate(transaction = if_else(transaction == "Imputed social contributions/Social benefits received" , "Social benefits received", transaction)) %>%
  mutate(transaction = if_else(transaction == "Social contributions/Social benefits paid" , "Social contributions, paid", transaction)) %>%
  mutate(transaction_code = case_when(transaction == "Operating surplus"                             ~ "Primary Resources",
                                      transaction == "Mixed income"                                  ~ "Primary Resources",
                                      transaction == "Compensation of employees"                     ~ "Primary Resources",
                                      transaction == "Property income, received"                     ~ "Primary Resources",
                                      transaction == "Property income, paid"                         ~ "Primary Uses",
                                      transaction == "Social benefits received"                      ~ "Secondary Resources",
                                      transaction == "Other current transfers, received"             ~ "Secondary Resources",
                                      transaction == "Current taxes on income, wealth etc"           ~ "Secondary Uses",
                                      transaction == "Social contributions, paid"                    ~ "Secondary Uses",
                                      transaction == "Other current transfers, paid"                 ~ "Secondary Uses",
                                      transaction == "Gross Disposable Income"                       ~ "Gross Disposable Income")) %>%
  mutate(source = nd3_flow$source[match(nd3_flow$source, gdhi_sankey_prep$transaction)]) %>%
  mutate(target = nd3_flow$target[match(gdhi_sankey_prep$transaction, nd3_names$names)]) %>%
  janitor::clean_names() %>%
  select(-nuts_level , -nuts_code) %>%
  filter(year == 2018 & region_name == "North East")


```

##### Example
```{r}

myDf <- 
list(
  nodes = data.frame(name   =            c( "A", "B", "C", "D", "E",
                                            "V", "W", "X", "Y", "Z")),
  links = data.frame(source = as.integer(c(0, 1, 2, 4, 3, 4, 4)),
                     target = as.integer(c(7, 6, 5, 8, 5, 8, 9)),
                     value  =            c(1, 4, 1, 5, 1, 5, 3))
)

sankeyNetwork(Links = myDf$links, Nodes = myDf$nodes, 
              Source = "source", Target = "target", Value = "value", 
              NodeID = "name", units = "TWh", fontSize = 25, nodeWidth = 30, 
              fontFamily = "sans-serif", iterations = 0, sinksRight = FALSE)

```

#### {networkD3}
```{r}
library(networkD3)

nd3_gdhi <- 
list(
  nodes = data.frame(name   = nd3_names$names),
  links = data.frame(source = as.integer(c(0,1,2,3,4,4,5,5,5,6,6,6,7, 8, 9)),
                     target = as.integer(c(4,4,4,4,5,6,7,8,9,7,8,9,10,10,10)),
                     value  = gdhi_sankey_prep$value))


sankeyNetwork(Links = nd3_gdhi$links, Nodes = nd3_gdhi$nodes ,
              Source = "source", Target = "target", Value = "value", NodeID = "name",
              units = "£", fontSize = 12, nodeWidth = 30, iterations = 0, sinksRight = FALSE)
```

year == 2018 & region_name == "London"

primary_resources, primary_uses, secondary_resources, secondary_uses, gross_disposable_income

region_name , operating_surplus, mixed_income, compensation_of_employees, property_income_received, property_income_paid,
            imputed_social_contributions_social_benefits_received, other_current_transfers_received, current_taxes_on_income_wealth_etc,
            social_contributions_social_benefits_paid, other_current_transfers_paid, gross_disposable_income
#### {ggsankey}
```{r}
library(ggsankey)
theme_set(theme_minimal())

# Make wide, make long, make Sankey all day long
long_nuts1 <-
make_long(wide_nuts1)

# plot ggsankey
nuts1_long %>%
  ggplot(aes(x = x, 
             next_x = next_x, 
             node = node, 
             next_node = next_node,
             fill = factor(node),
             label = node)) +
  geom_sankey(show.legend = FALSE) +
  geom_sankey_label(size = 3, color = "white", fill = "gray40") +
  theme_sankey(base_size = 18)

```


