---
title: "Per Head Regional Breakdowns"
author: "NearAndDistant"
date: "17/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Grab ONS Gross Data from GitHub (Script in Repository)

Gross Disposable household Income is the amount of money that that all of the individuals in the household sector have available for spending or saving after income distribution measures (for example, taxes, social contributions and benefits) have taken effect. GDHI does not provide measures relating to actual households or family units. The figures cover regions, sub-regions and local areas of the UK. It consists of:

##### Primary resources
0: Operating surplus
1: Mixed income
2: Compensation of employees
3: Property income, received

= Primary Resource total

##### Primary uses
4: Property income, paid

##### Secondary resources
5: Social benefits received
6: Other current transfers, received

= Balance of Primary Income

##### Secondary uses
7: Current taxes on income, wealth etc
8: Social contributions, paid
9: Other current transfers, paid

##### GDHI
= Gross Disposable Income

```{r}
library(tidyverse)

# you can see the primary script to grab this at https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gross_income_1997_to_2018/
ons_gdhi_components_perhead_raw <-
read_csv("https://raw.githubusercontent.com/NearAndDistant/ruk/main/data_scripts/ons_gdhi_97_18/ons_gdhi_perhead_components_97_18.csv")

```

##### Grabs NUTS1
```{r}

ons_gdhi_region <-
ons_gdhi_components_perhead_raw %>%
  janitor::clean_names(numerals = "left") %>%
  filter(nuts_level == "NUTS1") %>% # NUTS1 level is regions
  filter(region_name == "North East") %>% # pick region
  # filter(region_name == "North East" | region_name == "London") %>% 
  pivot_longer(cols = starts_with("x") , names_to = "year" , values_to = "value") %>%
  mutate(year = str_remove(year , "x"),
         year = as.numeric(year)) %>%
  filter(year == 2018) %>% # pick year
  # clean long names
  mutate(region_name = str_to_title(region_name)) %>%
  # mutate(region_name = if_else(region_name == "Yorkshire And The Humber", "Yorkshire", region_name)) %>%
  mutate(transaction = if_else(transaction == "Imputed social contributions/Social benefits received" , "Social benefits received", transaction)) %>%
  mutate(transaction = if_else(transaction == "Social contributions/Social benefits paid" , "Social contributions, paid", transaction)) %>%
  # filter out unecessary transactions i.e. totals or balances
  mutate(transaction_code = if_else(transaction == "Primary resources total" , "PRT", transaction_code)) %>% # so we keep this as a flow node
  mutate(value = if_else(transaction == "Gross Disposable Income" , NA_real_, value)) %>% # so we keep this as a flow node
  drop_na(transaction_code) %>% # NAs are just total rows in the ONS Excel file, unecessary here
  select(-year, -nuts_level, -nuts_code, -transaction_code) %>%
  # balancing out financials: as the flow diagram needs to know what is left over after "secondary payments" are made (taxes etc)
  # we have to calculate their subtraction from Balance of Payments so that Gross Disposable can be shown
  pivot_wider(id_cols = c(region_name:transaction), names_from = "transaction", values_from = "value") %>%
  mutate(`Balance of primary incomes` = `Primary resources total` - `Property income, paid` + `Social benefits received` + `Other current transfers, received`) %>%
  rename(`Secondary resources total` = `Balance of primary incomes`) %>% # aligning to the ONS figures on Secondary Income
  #   mutate(value = if_else(transaction == "Gross Disposable Income" , NA_real_, value)) %>% # so we keep this as a flow node
  pivot_longer(cols = c(`Operating surplus`:`Gross Disposable Income`), names_to = "transaction" , values_to = "value")

```

##### Adding Link Groups
```{r}
prep_region <-
ons_gdhi_region %>%
mutate(
source = 
case_when(
transaction == "Operating surplus"                   ~ "Operating surplus",
transaction == "Mixed income"                        ~ "Mixed income",
transaction == "Compensation of employees"           ~ "Compensation of employees",
transaction == "Property income, received"           ~ "Property income, received",
transaction == "Primary resources total"             ~ "Primary resources total",
transaction == "Property income, paid"               ~ "Primary resources total",
transaction == "Secondary resources total"           ~ "Secondary resources total",
transaction == "Social benefits received"            ~ "Social benefits received",
transaction == "Other current transfers, received"   ~ "Other current transfers, received",
transaction == "Current taxes on income, wealth etc" ~ "Secondary resources total",
transaction == "Social contributions, paid"          ~ "Secondary resources total",
transaction == "Other current transfers, paid"       ~ "Secondary resources total",
transaction == "Gross Disposable Income"             ~ "Secondary resources total"),
target = 
case_when(
transaction  == "Operating surplus"                   ~ "Primary resources total",
transaction  == "Mixed income"                        ~ "Primary resources total",
transaction  == "Compensation of employees"           ~ "Primary resources total",
transaction  == "Property income, received"           ~ "Primary resources total",
transaction  == "Primary resources total"             ~ "Secondary resources total",
transaction  == "Property income, paid"               ~ "Property income, paid",
transaction  == "Secondary resources total"          ~ "Gross Disposable Income",
transaction  == "Social benefits received"            ~ "Secondary resources total",
transaction  == "Other current transfers, received"   ~ "Secondary resources total",
transaction  == "Current taxes on income, wealth etc" ~ "Current taxes on income, wealth etc",
transaction  == "Social contributions, paid"          ~ "Social contributions, paid",
transaction  == "Other current transfers, paid"       ~ "Other current transfers, paid",
transaction  == "Gross Disposable Income"             ~ "Gross Disposable Income")) %>%
select(region_name , source , target , value) %>%
group_by(region_name)

# create line ids
indexed_region <-
prep_region %>%
  mutate(id = str_pad(7:19, width = 2, pad = "0")) %>%
  select(id , everything()) %>%
  ungroup()

```

```{r}

# create four new rows for the four income streams so we have a source for their sankey flows
ne_join <- 
  tibble(
    id          = indexed_region$id,
    region_name = indexed_region$region_name,
    source      = indexed_region$region_name,
    target      = indexed_region$source,
    value       = indexed_region$value) %>%
    filter(id %in% c("07", "08" , "09" , "10" , "14" , "15")) %>%
    # create new ids for 01 - 06
    mutate(id = str_pad(1:6, width = 2, pad = "0")) %>%
    select(id , everything())

# join 
region_full <-
  rbind(indexed_region , ne_join) %>%
  arrange(id)

```

#### Create networkD3 links

Great reference on this by NHS-R (Lisa Cummmins), here:
https://www.youtube.com/watch?v=3OvsW8OI1wo

```{r}

# nodes ne
nodes_ne <- 
  region_full %>%
  select(source , target, value) %>%
  pivot_longer(cols = c(source,target), values_to = "table_link") %>%
  select(-name) %>%
  distinct(table_link, .keep_all = TRUE) %>%
  mutate(value = case_when(table_link == "Primary resources total"   ~ 17818,
                           table_link == "Secondary resources total" ~ 24195,
                           table_link == "Gross Disposable Income"   ~ 16995,
                           TRUE ~ value)) %>%
  mutate(table_link_lab = c(paste0(table_link," (",scales::comma(value, prefix = "£"),")")),
         table_link_lab = ifelse(table_link == "North East", "North East, 2018", table_link_lab),
         table_link_lab = ifelse(table_link == "Current taxes on income, wealth etc", 
                                 paste0("Current taxes, paid (",scales::comma(value, prefix = "£"),")"), table_link_lab)) %>%
  select(-value)

```

#### Create node links - North East Region
```{r}

networkD3_table_ne <-
  region_full %>%
  mutate(source_id  = match(source , nodes_ne$table_link)-1) %>%
  mutate(target_id = match(target ,  nodes_ne$table_link)-1) %>%
  mutate(color_group = "Group")

link_color_region <- 
JS('d3.scaleOrdinal()
    .domain(["White", "Group"])
    .range(["#FFFFFF", "#529Ecc"])') # #00306FFF

```

#### {networkD3} Region
```{r}
library(networkD3)

nd3_gdhi_ne <- 
list(
  nodes = data.frame(name   =        nodes_ne$table_link_lab,
                     color  =        "will pick first color i.e. white"),
  links = data.frame(source =        networkD3_table_ne$source_id,
                     target =        networkD3_table_ne$target_id,
                     value  =        networkD3_table_ne$value,
                     color  =        networkD3_table_ne$color_group))


gdhi_ne_sankey <-
sankeyNetwork(Links = nd3_gdhi_ne$links, Nodes = nd3_gdhi_ne$nodes , 
              Source = "source", Target = "target", Value = "value", 
              LinkGroup = "color", colourScale = link_color_region,
              NodeID = "name", NodeGroup = "color", nodeWidth = 30,
              units = "(£m)", fontSize = 16, fontFamily = "Arial" ,
              width = 1700 , height = 900 ,  sinksRight = FALSE, iterations = 0,
              margin = list(right = 110, left = -20))

# render with JavaScript to change characteristics (#858b97)
# render with js
gdhi_ne_sankey_js <- 
htmlwidgets::onRender(gdhi_ne_sankey,
  'function(el, x) {
    d3.selectAll(".node text").attr("fill", "black");
    d3.selectAll(".node text").attr("text-anchor", "end").attr("x", -10);
    d3.select("body").style("background-color", "#FFFFFF"); 
  }'
) # #818380

# Shadow: #373737

gdhi_ne_sankey_js

```

#### Saving Image
```{r}

# create directory
dir.create("outputs")

dir.create("outputs/html")
dir.create("outputs/images")

# save it as html
saveNetwork(gdhi_ne_sankey_js, here::here("outputs/html/gdhi_region_sankey.html"))

library(webshot2)
# you convert it as png
webshot(here::here("outputs/html/gdhi_region_sankey.html"),
        here::here("outputs/images/gdhi_region_sankey.png"),
        vwidth = 1500, vheight = 800)

```

